---
title: "Surv 617 Group Project"
author: "Alex Caramanico"
date: "`r Sys.Date()`"
output: pdf_document
---
# Overview
## The Survey
### Size

* excluded institutional populations, such as persons in hospitals, hotels,
barracks camps, hostels, and prisons
* 629 clusters: 211 urban, 418 rural
* 26 households from each cluster (ended up a bit short of expected total)
* 16,312 households selected
* 15,907 occupied 
* 15,705 interviewed, overall response rate = 99%. 

Aged 15-49 are eligible for interview (*these files not in report*) \\

* In the interviewed households, 15,699 women 15,254 completed, response rate = 97%. 
* In the 50% of households of households selected for the male questionnaire, 
6,367 men 5,763 completed, response rate = 91%.

### Questionnaires:
-Household: \\
  age, sex, marital status, education, relationship to household head,
  Parents’ survival status for <18, disability status for 5+
  source of drinking water, type of toilet facilities, floor materials
  ownership of various durable goods, and ownership and use of mosquito nets.
-Woman’s: (*this and all following files not used in report*) \\
  age, education, media exposure, Birth history and childhood mortality,
  Knowledge and use of family planning methods, Fertility preferences (antenatal, 
  delivery, and postnatal care),Breastfeeding and infant feeding practices,
  Vaccinations and childhood illnesses, Marriage and sexual activity, Women’s 
  work and husband’s background characteristics, Other health issues, Adult 
  mortality,including maternal mortality, Female genital cutting, Early childhood 
  development, Malaria, Domestic violence
-Man’s: \\
  similar to womens, less reproductive info
-Biomarker: \\
  height and weight, anaemia tests, malaria tests, request for urine tests
-Micronutrient: \\
  anthropometric measurements, anaemia and malaria test results, and haematocrit 
  results for women and to document the outcome of venous blood collection 
  procedures and, in the relevant subset, the time for each step in the vitamin 
  A relative dose response test.
-Fieldworkers:
  self-administered, basic info (people working the survey)
  
## The Project
### Resources

* Data, Documentation: 
https://drive.google.com/drive/u/1/folders/1pE95hLbjKLqq7TxFdXj0HwwAcqCx7iIz
* Project Report:
https://www.overleaf.com/project/690d456c18e5aa3efcbdd9a4
* Google Doc
https://docs.google.com/document/d/1ts9Y1K_-uJ0UG1LI4pLNA6WkHsEVZklJihBx-RPYI3M/edit?tab=t.0
* GitHub
https://github.com/acforrester/SURV_621_Project
* Variable Codes
file:///C:/Users/jacma/OneDrive/school%20work/UMD/Surv%20617/Group%20Project/Variable%20Codes.pdf

### Goals/Ideas:

* Use the wealth index variable (constructed using PCA of household assets and 
  amenities) as outcome variable, see if wealth is clustered across regions 
  and/or place of residence (urban/rural/semi-urban) 
* Tanzania case study, ideally would be able to apply to other developing countries
* Policy implications: 
  * evidence of significant wealth clustering --> advise govts to target 
    interventions/policies that address regional inequalities.
  * no evidence of clustering, then the wealth differences could be more driven 
    by household level (L1) factors so policymakers can focus on improving 
    resource use and resource distributions at the household level.
* See if model can be used to improve district-level wealth estimates from the DHS. 
  Using covariates such as size and composition, rural/urban, and labor force 
  status --> unit-level small area model could produce estimates
  
# Coding 
## Load Packages
```{r,warning=F,message=F}
library(tidyverse)
library(haven)

library(ggplot2)
library(ggsurvey)
library(cowplot)

library(lme4)
library(glmnet)
library(rstatix) # for Cramer's V
library(Hmisc) # for weighted std deviation
library(survey)
library(srvyr)
library(stargazer) # for examining output
```

## Data Loading and Cleaning
```{r,warning=F,message=F}
tanzania_1 <- read_dta("C:/Users/jacma/OneDrive/school work/UMD/Surv 617/Group Project/TZHR81FL.DTA",
                       col_select = c(
                         hv001,   # cluster number (PSU)
                         hv002,   # household number
                         hv005,   # household sample weight
                         hv009,   # number of usual members 
                         hv022,   # sampling strata
                         hv024,   # region
                         hv025,   # type of place of residence (urban/rural)
                         hv201,   # source of drinking water (99:missing)
                         hv205,   # type of toilet facility (99:missing)
                         hv206,   # sharing of toilet (*this is electricity (9:missing)*)
                         hv213,   # main floor material (*hv208=tv*) (99:missing)
                         hv217,   # relationship structure
                         hv219,   # sex of household head (9:missing)
                         hv220,   # age of household head (9:missing)
                         hv221,   # phone (*hv246 is actually livestock*) (9:missing)
                         hv226,   # type of cooking fuel (99:missing)
                         hv247,   # has bank account (9:missing)
                         hv270,   # wealth index (combined)
                         hv271,   # wealth index (continuous)
                         hv270a,  # wealth index (rural/urban adjusted)
                         shreg2, # mainland/Zanzibar
                         hv106_01)# education of household head (9:missing)
) %>% mutate(hv005 = hv005 / 1e6)  

colnames(tanzania_1) <- c("ea","house_id","hh_weight","num_members","sampling_strata",
                          "region","residence","water","toilet","electric","floor",
                          "hh_struct","hh_sex","hh_age","phone","cooking","bank",
                          "wealth_ind_q","wealth_ind","wealth_ind_adj","govt",
                          "hh_educ")
tanzania_1 <- tanzania_1[,-2]

# grouping of categorical variables (recode water, toilet to safe/unsafe)
tanzania_1$water_improved <- case_when(
  tanzania_1$water %in% c(11, 12, 13, 14, 21, 31, 41, 51, 61, 62, 71) ~ "Improved",
  tanzania_1$water %in% c(32, 42, 43, 96)                             ~ "Unimproved")
  # where do 72, 73 go? "satchet water" - 10 homes, "water kiosk/domestic points" - 12 homes
  # 10,20,30,40 not actually in data 
tanzania_1$toilet_improved <- case_when(
  tanzania_1$toilet %in% c(11, 12, 13, 14, 15, 21, 22, 23) ~ "Improved",
  tanzania_1$toilet %in% c(31, 41, 42, 43, 95, 96) ~ "Unimproved")
  # where does 24 go? "pit latrine with slab (not washable)" - 3423 homes
  # 10,20,30 not actually in data

# fix factor names
vars <- c("hh_educ", "water", "toilet", "floor", "cooking","wealth_ind_q", 
          "wealth_ind_adj", "residence","region", "hh_struct", "govt")
for (v in vars) {
  lbls <- attr(tanzania_1[[v]], "labels")
  tanzania_1[[v]] <- factor(tanzania_1[[v]],
                            levels = lbls,
                            labels = names(lbls))
}
rm(vars,lbls,v)

# any renaming of factor levels goes here (basically just for improving visuals/tables)
levels(tanzania_1$hh_educ) <- c("none/early","primary","secondary","higher", "don't know")

# calculated / engineered features
ages <- read_dta("C:/Users/jacma/OneDrive/school work/UMD/Surv 617/Group Project/TZHR81FL.DTA",
                 col_select =c(hv105_01:hv105_56))
tanzania_1$dep_ratio <- apply(ages,1,
                              function(row){
                                100 * sum(row %in% c(0:14,65:99))/sum(row %in% 15:64)
                                })
rm(ages) 

tanzania_1$dep_ratio2 <- tanzania_1$dep_ratio
tanzania_1$dep_ratio <- pmin(tanzania_1$dep_ratio, 100) # fix dividing by 0 issue
# this is weird...only 723 are dividing by 0. 5761 are being truncated to 100
tanzania_1$dep_ratio2[tanzania_1$dep_ratio2==Inf] <- NA # do this instead?

# survey design
DHSdesign <- tanzania_1 %>% as_survey_design(ids=ea,
                                        weights=hh_weight,
                                        strata=sampling_strata,
                                        nest=TRUE)
```

## Missing Data Analysis 
```{r,warning=F,message=F}
apply(tanzania_1,2,function(col){sum(is.na(col))}) # NAs only for variable recodes

MISSING = FALSE
if (MISSING){
  # missing may be encoded as 9 or 99 (*found none*)
  apply(tanzania_1[,c(5,6,7,13,14,19)],2,function(col){sum(col==9)})
  apply(tanzania_1[,c(8,9,11,12)],2,function(col){sum(col==99)})
  library(visdat)
}
rm(MISSING)

### compare rows with NA to remaining data
tanzania_1$missing_info <- case_when(
  !(tanzania_1$water_improved %>% is.na()) & 
    !(tanzania_1$toilet_improved %>% is.na()) ~ "Neither", # 12262
  tanzania_1$water_improved %>% is.na() & 
    tanzania_1$toilet_improved %>% is.na() ~ "Both", # only 2
  tanzania_1$water_improved %>% is.na() ~ "Water Quality", # only 20
  tanzania_1$toilet_improved %>% is.na() ~ "Toilet Quality") # 3421
#tanzania_1$missing_info %>% table()

tanzania_1$nan_dep <- ifelse(tanzania_1$dep_ratio2 %>% is.na(), "Yes","No")

PLOT_MISSING <- TRUE
if (PLOT_MISSING){
  ### compare dependent variable across missingness state
  dependent <- list()
  dependent[[1]] <- ggplot(tanzania_1) +
    geom_density(mapping=aes(x=wealth_ind,group=missing_info,fill=missing_info),
                 alpha=0.4) +
    labs(title="Distribution of wealth index according to missingness")
  # definitely some differences here, but consider sample size
  
  dependent[[2]] <- ggplot(tanzania_1) +
    geom_boxplot(mapping=aes(y=wealth_ind,group=govt,fill=govt)) +
    theme(axis.text.x = element_blank(),axis.ticks.x = element_blank()) +
    facet_wrap(~missing_info) + 
    labs(title="Dispersion of Wealth Index across govt regions",
         subtitle="based on which variables are missing")
  # no clear differences

  dependent[[3]] <- ggplot(tanzania_1) +
    geom_boxplot(mapping=aes(y=wealth_ind,group=region,fill=region)) +
    theme(axis.text.x = element_blank(),axis.ticks.x = element_blank()) +
    facet_wrap(~missing_info) + 
    labs(title="Dispersion of Wealth Index across named regions",
         subtitle="based on which variables are missing")
  # no clear differences
  
  dependent[[4]] <- ggplot(tanzania_1) +
    geom_boxplot(mapping=aes(y=wealth_ind,group=region,fill=region)) +
    theme(axis.text.x = element_blank(),axis.ticks.x = element_blank()) +
    facet_wrap(~nan_dep) + 
    labs(title="Dispersion of Wealth Index across named regions",
         subtitle="Dependency Ratio 2 Missing or Not") 
  # much narrower range for missing, to be expected
  
  ### compare spread of cts variables
  independent <- list()
  independent[[1]] <- ggplot(tanzania_1) +
    geom_histogram(mapping=aes(x=num_members),bins=30,stat='count') +
    facet_wrap(~missing_info) + 
    labs(title="Distribution of # Members",
         subtitle="based on which variables are missing")
  # no clear differences, low sample sizes for "both" and "neither"

  independent[[2]] <- ggplot(tanzania_1) +
    geom_histogram(mapping=aes(x=hh_age),bins=30) +
    facet_wrap(~missing_info) + 
    labs(title="Distribution of household head age")
  # household head age may be a little more flatly distributed when toilet_improved
  # is missing, but I don't think there's anything intuitive here
  
  independent[[3]] <- ggplot(tanzania_1) +
    geom_histogram(mapping=aes(x=dep_ratio),bins=30) +
    facet_wrap(~missing_info) + 
    labs(title="Distribution of dependency ratio")
  # not obvious, this dependency ratio calculation may be problematic as well
  
  independent[[4]] <- ggplot(tanzania_1) +
    geom_histogram(mapping=aes(x=num_members),bins=30,stat='count') +
    facet_wrap(~nan_dep) + 
    labs(title="Distribution of # Members",
         subtitle="Dependency ratio NA?")
  # pretty similar shape, again sample size

  independent[[5]] <- ggplot(tanzania_1) +
    geom_histogram(mapping=aes(x=hh_age),bins=30) +
    facet_wrap(~nan_dep) + 
    labs(title="Distribution of household head age",
         subtitle="Dependency ratio NA?")
  # **very clear difference, missing dep_ratio2 may be related to older hh heads**
  
  missing_plots <- list(Dependent_Var=dependent,
                        Independent_Vars=independent)
}
if (exists("missing_plots")) {
  print(missing_plots)
  rm(missing_plots,dependent,independent)} 

tanzania_1$missing_info <- NULL
tanzania_1$nan_dep <- NULL

rm(MISSING,PLOT_MISSING)
```

## Exploratory Data Analysis
Variables in data frame but not explored in detail in this section include the
following, where (O) means they were included in some old analysis: phone, 
electric, wealth_ind_adj, wealth_ind_q (O). There are of course *many* other
variables which may be of use in the DHS, both within the household file and by 
compiling individual-level information (IR, MR) in a manner similar to how 
dependency ratio was calculated. 
### Statistics
#### Categorical Variables
```{r,warning=F,error=F,message=F}
apply(tanzania_1[,c(5:12,14:17,19:23)],2,table) # tables should sum to 15705 (unique houses)

## Cross-tabs to look at relationship between wealth quintile and categorical variables
xtabs(~region+wealth_ind_q,data=tanzania_1) # %>% heatmap(Colv=NA,Rowv=NA)
xtabs(~govt+wealth_ind_q,data=tanzania_1) # %>% heatmap(Colv=NA,Rowv=NA)
xtabs(~water+wealth_ind_q,data=tanzania_1)# heatmap(Colv=NA,Rowv=NA)
xtabs(~hh_struct+wealth_ind_q,data=tanzania_1) # %>% heatmap(Colv=NA,Rowv=NA)
xtabs(~toilet+wealth_ind_q,data=tanzania_1)# %>% heatmap(Colv=NA,Rowv=NA)
xtabs(~cooking+wealth_ind_q,data=tanzania_1)# %>% heatmap(Colv=NA,Rowv=NA)
xtabs(~bank+wealth_ind_q,data=tanzania_1)# %>% heatmap(Colv=NA,Rowv=NA)
xtabs(~residence+wealth_ind_q,data=tanzania_1)# %>% heatmap(Colv=NA,Rowv=NA)

xtabs(~wealth_ind_adj+wealth_ind_q,tanzania_1) # see where the adjustments were made, most people were moved into higher wealth brackets. 'middle' was essentially gutted

cramer_v(tanzania_1$wealth_ind_q,tanzania_1$wealth_ind_adj) # suspected collinearity
cramer_v(tanzania_1$electric,tanzania_1$bank) # suspected collinearity
cramer_v(tanzania_1$toilet,tanzania_1$water)
cramer_v(tanzania_1$toilet,tanzania_1$cooking)
cramer_v(tanzania_1$water,tanzania_1$cooking)

cramer_v(tanzania_1$water_improved,tanzania_1$toilet_improved) 
# ^ suspected collinearity, but actually not much
```

#### Continuous Variables
```{r}
## summarize cts variables - means, var, by region/other categories
summary(tanzania_1[,c(3,13,18,24,25)]) # unweighted summary
weighted_summary <- DHSdesign %>%
  dplyr::summarize(
    num_members = survey_mean(num_members, na.rm = TRUE, vartype = "se"), # or "var"?
    hh_age = survey_mean(hh_age, na.rm = TRUE, vartype = "se"),
    wealth_ind = survey_mean(wealth_ind, na.rm = TRUE, vartype = "se"),
    dep_ratio = survey_mean(dep_ratio, na.rm = TRUE, vartype = "se"),
    dep_ratio2 = survey_mean(dep_ratio2, na.rm = TRUE, vartype = "se")
  )
rbind(Min=cbind(min(tanzania_1$num_members),min(tanzania_1$hh_age),
                min(tanzania_1$wealth_ind),min(tanzania_1$dep_ratio),
                min(tanzania_1$dep_ratio2,na.rm=T)),
      Mean=weighted_summary[1,c(1,3,5,7,9)] %>% unlist() %>% round(3),
      Max=cbind(max(tanzania_1$num_members),max(tanzania_1$hh_age),
                max(tanzania_1$wealth_ind),max(tanzania_1$dep_ratio),
                max(tanzania_1$dep_ratio2,na.rm=T)),
      StdErr=weighted_summary[c(2,4,6,8,10)] %>% unlist() %>% round(3))
rm(weighted_summary)

regional_summary <- tanzania_1 %>% # summary statistics by region, unweighted
  group_by(region) %>%
  dplyr::summarize(
    mean_wealth = mean(wealth_ind),
    std_wealth = sd(wealth_ind),
    mean_members=mean(num_members),
    mean_head_age=mean(hh_age),
    mean_dep_ratio=mean(dep_ratio),
    mean_dep_ratio2=mean(dep_ratio2,na.rm=T)) %>% 
  arrange(desc(mean_wealth))
regional_summary %>% View()

regional_summary_wt <- tanzania_1 %>% # summary statistics by region, weighted
  group_by(region) %>%
  dplyr::summarize(
    mean_wealth = weighted.mean(wealth_ind,hh_weight),
    std_wealth = wtd.var(wealth_ind,hh_weight) %>% sqrt(),
    mean_members=weighted.mean(num_members,hh_weight),
    mean_head_age=weighted.mean(hh_age,hh_weight),
    mean_dep_ratio=weighted.mean(dep_ratio),
    mean_dep_ratio2=weighted.mean(dep_ratio2,na.rm=T)) %>% 
  arrange(desc(mean_wealth))
regional_summary_wt %>% View()

govt_summary <- tanzania_1 %>% # summary statistics by govt region, unweighted
  group_by(govt) %>%
  dplyr::summarize(
    mean_wealth = mean(wealth_ind),
    std_wealth = sd(wealth_ind),
    mean_members=mean(num_members),
    mean_head_age=mean(hh_age),
    mean_dep_ratio=mean(dep_ratio),
    mean_dep_ratio2=mean(dep_ratio2,na.rm=T)) %>% 
  arrange(desc(mean_wealth))
govt_summary %>% View()

govt_summary_wt <- tanzania_1 %>% # summary statistics by govt region, weighted
  group_by(govt) %>%
  dplyr::summarize(
    mean_wealth = weighted.mean(wealth_ind,hh_weight),
    std_wealth = wtd.var(wealth_ind,hh_weight) %>% sqrt(),
    mean_members=weighted.mean(num_members,hh_weight),
    mean_head_age=weighted.mean(hh_age,hh_weight),
    mean_dep_ratio=weighted.mean(dep_ratio),
    mean_dep_ratio2=weighted.mean(dep_ratio2,na.rm=T)) %>% 
  arrange(desc(mean_wealth))
govt_summary_wt %>% View() 
# could also do the above grouping by other categorical variables - visualizations do this
```

### Visualizations (to be saved in single object where possible)
#### Univariate
```{r}
# distributions of continuous random variables
univariate <- plot_grid(ggplot(tanzania_1) +
                          geom_histogram(mapping=aes(x=num_members),bins=30) +
                          labs(title="Distribution of # Members"),
                        ggplot(tanzania_1) +
                          geom_histogram(mapping=aes(x=hh_age),bins=40) +
                          labs(title="Distribution of Household-Head Age"),
                        ggplot(tanzania_1) +
                          geom_histogram(mapping=aes(x=dep_ratio),bins=30) +
                          labs(title="Distribution of Dependency Ratio"),
                        ggplot(tanzania_1) +
                          geom_histogram(mapping=aes(x=dep_ratio2),bins=30) +
                          labs(title="Distribution of Dependency Ratio 2"),
                        ggplot(tanzania_1) +
                          geom_histogram(mapping=aes(x=wealth_ind),bins=60) +
                          labs(title="Distribution of Wealth Index (cts)",
                               x="Wealth Index",
                               y="Count") +
                          theme(axis.text = element_text(size = 12)),nrow=3)
```

#### Dispersion of Wealth - Location
```{r,warning=F}
# how the spread of wealth varies according to location (regional variables)
location <- list()
location[[1]] <- ggplot(tanzania_1) +
  geom_boxplot(mapping=aes(y=wealth_ind,group=govt,fill=govt)) +
  theme(axis.text.x = element_blank(),axis.ticks.x = element_blank()) +
  labs(title="Dispersion of Wealth Index across govt regions",y="Wealth Index")

location[[2]] <- ggplot(tanzania_1) +
  geom_density(mapping=aes(x=wealth_ind,fill=govt),alpha=0.4) +
  labs(title="Distribution of Wealth Index (cts)")

location[[3]] <- ggplot(tanzania_1) +
  geom_boxplot(mapping=aes(y=wealth_ind,group=govt,fill=govt)) +
  facet_wrap(~bank) + # or electric, or cluttered water, toilet, cooking 
  theme(axis.text.x = element_blank(),axis.ticks.x = element_blank()) +
  labs(title="Dispersion of Wealth Index across govt regions",subtitle="bank or no bank",
       y="Wealth Index")

location[[4]] <- ggplot(tanzania_1) +
  geom_boxplot(mapping=aes(y=wealth_ind,group=region,fill=region)) +
  theme(axis.text.x = element_blank(),axis.ticks.x = element_blank()) +
  labs(title="Dispersion of Wealth Index across named regions",y="Wealth Index")

location[[5]] <- ggplot(tanzania_1) +
  geom_density(mapping=aes(x=wealth_ind,fill=residence),alpha=0.4) +
  labs(title="Distribution of Wealth Index (cts)",subtitle="by residence type",y="Wealth Index")
```

#### Dispersion of Wealth - Household Facilities
```{r,warning=F}
# how the distribution of wealth varies according to household facilities and amenities 
# (e.g. water, toilet, banking)
facilities <- list()
facilities[[1]] <- ggplot(tanzania_1) +
  geom_boxplot(mapping=aes(y=wealth_ind,group=water,fill=water)) +
  theme(axis.text.x = element_blank(),axis.ticks.x = element_blank()) +
  labs(title="Dispersion of Wealth Index", subtitle="Across types of water sources",y="Wealth Index")

facilities[[2]] <-plot_grid(ggplot(tanzania_1 %>% na.omit()) +
                              geom_boxplot(mapping=aes(y=wealth_ind,
                                                       group=water_improved,
                                                       fill=water_improved)) +
                              theme(axis.text.x = element_blank(),
                                    axis.ticks.x = element_blank(),
                                    legend.position="none") +
                              labs(title="Dispersion of wealth",
                                   subtitle="by water quality",y="Wealth Index"),
                            ggplot(tanzania_1 %>% na.omit()) +
                              geom_density(mapping=aes(x=wealth_ind,
                                                       group=water_improved,
                                                       fill=water_improved),alpha=0.4) +
                              theme(axis.text.x = element_blank(),
                                    axis.ticks.x = element_blank()) +
                              labs(title="Distribution of wealth",
                                   subtitle="by water quality",y="Wealth Index"))
facilities[[3]] <- ggplot(tanzania_1) +
  geom_boxplot(mapping=aes(y=wealth_ind,group=cooking,fill=cooking)) +
  theme(axis.text.x = element_blank(),axis.ticks.x = element_blank()) +
  labs(title="Dispersion of Wealth Index",
       subtitle="Across types of cooking fuel",y="Wealth Index")

facilities[[4]] <- ggplot(tanzania_1) +
  geom_boxplot(mapping=aes(y=wealth_ind,group=toilet,fill=toilet)) +
  theme(axis.text.x = element_blank(),axis.ticks.x = element_blank()) +
  labs(title="Dispersion of Wealth Index",
       subtitle="Across toilet types",y="Wealth Index")

facilities[[5]] <-plot_grid(ggplot(tanzania_1 %>% na.omit()) +
                              geom_boxplot(mapping=aes(y=wealth_ind,
                                                       group=toilet_improved,
                                                       fill=toilet_improved)) +
                              theme(axis.text.x = element_blank(),
                                    axis.ticks.x = element_blank(),
                                    legend.position="none") +
                              labs(title="Dispersion of wealth",
                                   subtitle="by toilet quality",x="Wealth Index"),
                            ggplot(tanzania_1 %>% na.omit()) +
                              geom_density(mapping=aes(x=wealth_ind,
                                                       group=toilet_improved,
                                                       fill=toilet_improved),alpha=0.4) +
                              theme(axis.text.x = element_blank(),
                                    axis.ticks.x = element_blank()) +
                              labs(title="Distribution of wealth",
                                   subtitle="by toilet quality",x="Wealth Index"))

facilities[[6]] <- ggplot(tanzania_1) +
  geom_density(mapping=aes(x=wealth_ind,fill=factor(bank)),alpha=0.4) +
  scale_fill_discrete(labels=c("No","Yes")) + 
  labs(title="Distribution of Wealth Index (cts)",subtitle="Bank account or not",
       x="Wealth Index",fill="bank account")
```

#### Dispersion of Wealth - Household Characteristics
```{r,warning=F}
# how the distribution of wealth varies according to characteristics of household, 
# e.g. sex and education of household head, relationship structure
characteristics <- list()
characteristics[[1]] <- ggplot(tanzania_1) +
  geom_density(mapping=aes(x=wealth_ind,group=factor(hh_sex),fill=factor(hh_sex)),alpha=0.4) +
  scale_fill_discrete(labels=c("Male","Female")) +
  labs(title="Dispersion of Wealth Index by sex of household head",
       x="Wealth Index",fill="sex")

characteristics[[2]] <- ggplot(tanzania_1) +
  geom_boxplot(mapping=aes(y=wealth_ind,group=hh_educ,fill=hh_educ)) +
  theme(axis.text.x = element_blank(),axis.ticks.x = element_blank()) +
  labs(title="Dispersion of Wealth Index by education level of household head",
       y="Wealth Index",fill="Highest Education")

characteristics[[3]] <- plot_grid(ggplot(tanzania_1) +
                                    geom_density(mapping=aes(x=wealth_ind,fill=hh_educ),
                                                 alpha=0.4) +
                                    guides(fill="none") + 
                                    labs(title="Distribution of Wealth Index (cts)",
                                         subtitle="by education of HH head",
                                         x="Wealth Index"),
                                  ggplot(tanzania_1) +
                                    geom_histogram(mapping=aes(x=wealth_ind,fill=hh_educ),
                                                   bins=30) +
                                    labs(title="Distribution of Wealth Index (cts)",
                                         subtitle="by education of HH head",
                                         x="Wealth Index",fill="Highest Education"))

characteristics[[4]] <- ggplot(tanzania_1) +
  geom_boxplot(mapping=aes(y=wealth_ind,group=hh_struct,fill=hh_struct)) +
  theme(axis.text.x = element_blank(),axis.ticks.x = element_blank()) +
  labs(title="Dispersion of Wealth Index across relationship structures")

characteristics[[5]] <- ggplot(tanzania_1) +
  geom_density(mapping=aes(x=wealth_ind,fill=hh_struct),alpha=0.4) +
  labs(title="Distribution of Wealth Index (cts)",
       subtitle="by relationship structure") # no adults only has 2 observations


wealth_dispersion_cat <- list(location,facilities,characteristics)
rm(location,facilities,characteristics)
```

#### Continuous Predictors - Aggregated and Non-Aggregated
```{r,warning=F}
# relationship between wealth index and continuous predictors
### non-aggregated
wealth_dispersion_cts <- list()
wealth_dispersion_cts[[1]] <- ggplot(tanzania_1) +
  geom_point(mapping=aes(x=num_members,y=wealth_ind,color=factor(bank))) +
  scale_color_discrete(labels=c("No","Yes")) +
  labs(y="Wealth Index",x="Numer of Members",color="Has Bank Account")

wealth_dispersion_cts[[2]] <- ggplot(regional_summary) +
  geom_point(mapping=aes(x=mean_members,y=mean_wealth)) +
  theme(legend.position="none") +
  labs(title="Avg Members vs avg wealth index", subtitle="grouped by region",
       x="Avg members per house in region",y="Avg wealth index in region")

wealth_dispersion_cts[[3]] <- ggplot(tanzania_1) +
  geom_point(mapping=aes(x=dep_ratio,y=wealth_ind,color=factor(bank)),size=0.5) +
  theme(legend.position="none") +
  labs(title="Dependency ratio vs wealth index", subtitle="blue: has bank account")

wealth_dispersion_cts[[4]] <- ggplot(regional_summary) +
  geom_point(mapping=aes(x=mean_dep_ratio,y=mean_wealth)) +
  theme(legend.position="none") +
  labs(title="Dependency ratio vs wealth index", subtitle="grouped by region",
       x="Avg dependency ratio in region",y="Avg wealth index in region")

wealth_dispersion_cts[[5]] <- ggplot(regional_summary) +
  geom_point(mapping=aes(x=mean_dep_ratio2,y=mean_wealth)) +
  theme(legend.position="none") +
  labs(title="Dependency ratio (fixed?) vs wealth index", subtitle="grouped by region",
       x="Avg dependency ratio in region",y="Avg wealth index in region")

wealth_dispersion_cts[[6]] <- ggplot(tanzania_1) + # confetti
  geom_point(mapping=aes(x=hh_age,y=wealth_ind,color=region),size=0.2) +
  labs(title="Number of members vs wealth index", subtitle="colored by region")

### aggregated by region or govt
# pairs plots, cannot be saved into list
panel.smaller_points <- function(x, y, ...) {points(x, y, cex = 0.5, ...)}
pairs(tanzania_1[,c(3,4,13,18)],
      lower.panel = panel.smaller_points, 
      upper.panel = panel.smaller_points,)
pairs(regional_summary[,c(2,4:7)]) # no relationships with outcome variable aside from dep_ratio
pairs(govt_summary[,c(2,4:7)]) # not really any relationship, small "sample" anyway
rm(regional_summary,govt_summary,panel.smaller_points)
```

#### Exploration
```{r,warning=F}
all_ggs <- list(univariate,wealth_dispersion_cat,wealth_dispersion_cts)
rm(univariate,wealth_dispersion_cat,wealth_dispersion_cts)

all_ggs
# clear discrepancies in wealth across administrative/government regions, 
#   urban/rural, possession of a bank account, educational attainment of 
#   household head, dependency ratio
# no clear discrepanices between different household relationship structures, 
#   sex or age of household head, number of members in household

rm(all_ggs)
```

#### With Sampling Weights Applied
```{r}
### location
# administrative regions
ggboxweight2d_svy(DHSdesign,wealth_ind,region) +
  aes(fill=region) + labs(y="Wealth Index") +theme(axis.text.x=element_blank(),
                                                     axis.ticks.x=element_blank()) 
# urban/rural
ggboxweight2d_svy(DHSdesign,wealth_ind,residence) + # not what I intended but cool
  aes(fill=region) + labs(y="Wealth Index") + theme(legend.position="none")
ggboxweight2d_svy(DHSdesign,wealth_ind,residence) +
  aes(fill=residence) + labs(y="Wealth Index") + theme(legend.position="none")
# government region
ggboxweight2d_svy(DHSdesign,wealth_ind,govt) +
  aes(fill=govt) + labs(y="Wealth Index") + theme(legend.position="none")
gghistweight_svy(DHSdesign,wealth_ind,binwidth = 5000) +
  aes(fill=govt) + labs(x="Wealth Index",fill="Government Region") 
ggboxweight2d_svy(DHSdesign,wealth_ind,govt) + # cool additional
  aes(fill=factor(hh_sex)) + labs(y="Wealth Index",x="",fill="Sex") + scale_fill_discrete(labels=c("Male","Female"))
ggboxweight2d_svy(DHSdesign,wealth_ind,govt) + # cool additional
  aes(fill=hh_educ) + labs(y="Wealth Index",x="",fill="Highest Education")
gghistweight2d_svy(DHSdesign,wealth_ind,govt) + # cool additional
  aes(fill=hh_educ) + labs(x="Wealth Index",fill="Highest Education")
ggboxweight2d_svy(DHSdesign,wealth_ind,govt) + # cool additional
  aes(fill=factor(bank)) + labs(y="Wealth Index",x="",fill="Has bank account") + scale_fill_discrete(labels=c("No","Yes"))

### facilities
# water source
ggboxweight2d_svy(DHSdesign,wealth_ind,water) +
  aes(fill=water) + labs(y="Wealth Index",x="") + theme(axis.text.x=element_blank(),
                                                        axis.ticks.x=element_blank())
ggboxweight2d_svy(DHSdesign,wealth_ind,water_improved) +
  aes(fill=water_improved) + labs(y="Wealth Index",x="Water Source Quality") + theme(legend.position="none")
gghistweight2d_svy(DHSdesign,wealth_ind,water_improved) + 
  aes(fill=water_improved) + labs(x="Wealth Index") + theme(legend.position="none")

# toilet facility
ggboxweight2d_svy(DHSdesign,wealth_ind,toilet) +
  aes(fill=toilet) + labs(y="Wealth Index",x="") + theme(axis.text.x=element_blank(),
                                                         axis.ticks.x=element_blank())
ggboxweight2d_svy(DHSdesign,wealth_ind,toilet_improved) +
  aes(fill=toilet_improved) + labs(y="Wealth Index",x="Toilet Facility Quality") + theme(legend.position="none")
gghistweight2d_svy(DHSdesign,wealth_ind,toilet_improved) + 
  aes(fill=toilet_improved) + labs(x="Wealth Index") + theme(legend.position="none")

# flooring
ggboxweight2d_svy(DHSdesign,wealth_ind,floor) +
  aes(fill=floor) + labs(y="Wealth Index",x="") + theme(axis.text.x=element_blank(),
                                                     axis.ticks.x=element_blank())
# cooking
ggboxweight2d_svy(DHSdesign,wealth_ind,cooking) +
  aes(fill=cooking) + labs(y="Wealth Index",x="") + theme(axis.text.x=element_blank(),
                                                     axis.ticks.x=element_blank())

# bank account, electricity
ggboxweight2d_svy(DHSdesign,wealth_ind,factor(bank)) +
  aes(fill=factor(bank)) + labs(y="Wealth Index",x="Bank") + theme(legend.position="none")
ggboxweight2d_svy(DHSdesign,wealth_ind,factor(electric)) +
  aes(fill=factor(electric)) + labs(y="Wealth Index",x="Electricity") + theme(legend.position="none")

### characteristics
# gender
ggboxweight2d_svy(DHSdesign,wealth_ind,factor(hh_sex)) +
  aes(fill=factor(hh_sex)) + labs(y="Wealth Index",fill="Sex",x="") + scale_fill_discrete(labels=c("Male","Female"))
gghistweight2d_svy(DHSdesign,wealth_ind,factor(hh_sex)) + 
  aes(fill=factor(hh_sex)) + labs(x="Wealth Index",fill="Sex") + scale_fill_discrete(labels=c("Male","Female"))
# household structure
ggboxweight2d_svy(DHSdesign,wealth_ind,hh_struct) +
  aes(fill=factor(hh_struct)) + labs(y="Wealth Index",fill="Household Structure")
# education level
ggboxweight2d_svy(DHSdesign,wealth_ind,hh_educ) +
  aes(fill=hh_educ) + labs(fill="Highest Education",x='') + theme(axis.text.x=element_blank(),
                                                             axis.ticks.x=element_blank()) 
gghistweight_svy(DHSdesign,wealth_ind,binwidth = 5000) +
  aes(fill=hh_educ) + labs(x="Wealth Index") + theme(legend.position="none")
ggboxweight2d_svy(DHSdesign,wealth_ind,hh_educ) +
  aes(fill=factor(hh_sex)) + labs(y="Wealth Index",x="Education Level",fill="Sex") + 
  scale_fill_discrete(labels=c("Male","Female")) # I like this one

# pretty much all the same conclusions overall
```


## Modeling
Region and place of residence are *fixed* effects - all are included in the data
### Variable Transformations
```{r,echo=F}
# mean center num_members, age of household head at 0
tanzania_1$num_members <- tanzania_1$num_members - mean(tanzania_1$num_members)
tanzania_1$hh_age <- tanzania_1$hh_age - mean(tanzania_1$hh_age)

# scale wealth in 100,000s
tanzania_1$wealth_ind <- tanzania_1$wealth_ind /1e5

### CLUSTER = ea, REGIONID = region, URBAN = residence, ZANZIBAR = govt
### For 'education': none/early is the reference
### For 'region': dodoma is the reference
### For 'govt': mainland urban is the reference 
### For 'residence': urban is the reference
### For 'hh_sex': males are the reference
### --> so, intercept interpretations are for household with none/early 
#       education male as household head, located in dodoma or mainland urban or 
#       urban (depending upon which fixed effect used), with average number of 
#       members and no dependents (dep_ratio=0)
### insights are assumed to be the same across models unless *explicitly stated*
```

### Fitting Models
Variable selection ideas: Lasso, step-wise regression, random forest, VIF
```{r,echo=F}
## Model 1: FE for Region + Bank
mod1 <- lmer(wealth_ind ~
    factor(bank) + factor(hh_sex) + num_members + dep_ratio + factor(hh_educ)
    + factor(region) + (1 | ea), # random intercepts at household level
    data = tanzania_1,
    #weights = hh_weight,
    REML = T)
summary(mod1) # random intercepts seems logical - model deviations in wealth within EA, 
# overall intercept negative, significant
# bank EXTREMELY significant, increases wealth index
# sex of household head very insignificant (negative estimate though)
# household size very significant, positive estimate (higher increases wealth (?))
# dependency ratio negative and significant, makes sense (more dependents, less wealth)
# all levels of education significant, the higher the education the greater the effect on wealth
# some regions more significant as predictors than others: arusha, kilimanjaro, dar es salaam (!), 
#   kusini unguja, mjini magharibi (!) significantly wealthier. lindi, mtwara significantly poorer.
#   several regions have virtually no effect - ruvuma, mara, geita. these regions may have a mix
#   of urban/rural characteristics, small sample sizes, etc (see cross-tabs). they could also just
#   be very similar to the reference region dodoma

mod1b <- lmer(wealth_ind ~
    factor(bank) + factor(hh_sex) + num_members + dep_ratio2 + factor(hh_educ)
    + factor(region) + (1 | ea), # random intercepts at household level
    data = tanzania_1,
    #weights = hh_weight,
    REML = T)

## Model 2: FE for Tanzania + Bank
mod2 <- lmer(wealth_ind ~
    factor(bank) + factor(hh_sex) + num_members + dep_ratio + factor(hh_educ)
    + factor(govt) + (1 | ea), # random intercepts at household level
    data = tanzania_1,
    #weights = hh_weight,
    REML = T)
summary(mod2)
# intercept is positive now, and even more significant...
# *all* government regions are statistically significant, all decrease wealth
#   index compared to mainland urban (the baseline), pemba reduces the most, 
#   mainland rural the smallest reduction

mod2b <- lmer(wealth_ind ~
    factor(bank) + factor(hh_sex) + num_members + dep_ratio2 + factor(hh_educ)
    + factor(govt) + (1 | ea), # random intercepts at household level
    data = tanzania_1,
    #weights = hh_weight,
    REML = T)


## Model 3: FE for Zanzibar, RE for bank
mod3 <- lmer(wealth_ind ~
    factor(bank) + factor(hh_sex) + num_members + factor(hh_educ) + residence
    + factor(region) + (1 + factor(bank) | ea), # random intercepts at household level
  data = tanzania_1,
  #weights = hh_weight,
  REML = T)
summary(mod3)
# positive intercept estimate, less significant than model 2
# household sex now significant at 0.1, being female decreases wealth
# rural residences have *much* lower wealth, extremely signifcant and large 
#   negative parameter estimate
# arusha, kilimanjaro (!), dar es salaam (!), kusini unguja, mjini magharibi (!!!) significantly
#   wealthier; lindi, mtwara, tabora, rukwa, kigoma, manyara, simiyu significantly poorer; kagera,
#   mara, geita must be similar to dodoma

# Model 4: FE for governmental region, RE for education of household head
mod4 <- lmer(wealth_ind ~
    factor(bank) + factor(hh_sex) + num_members + factor(hh_educ) + factor(govt) + 
      (1 + factor(hh_educ) | ea), # random intercepts at household level
  data = tanzania_1,
  #weights = hh_weight,
  REML = T)
summary(mod4) # very small variance estimates on these - effect of education likely the same across EAs

# Model 5: FE for governmental region, RE for sex of household head
mod5 <- lmer(wealth_ind ~
    factor(bank) + factor(hh_sex) + num_members + factor(hh_educ) + factor(govt) + 
      (1 + factor(hh_sex) | ea), # random intercepts at household level
  data = tanzania_1,
  #weights = hh_weight,
  REML = T)
summary(mod5) # small variance, effect of gender the same across regions

# Model 6: FE for governmental region, RE for number of members
mod6 <- lmer(wealth_ind ~
    factor(bank) + factor(hh_sex) + num_members + factor(hh_educ) + factor(govt) + 
      (1 + num_members | ea), # random intercepts at household level
  data = tanzania_1,
  #weights = hh_weight,
  REML = T)
summary(mod6) # INCREDIBLY small variance, effect of household size same across regions

rm(mod1,mod1b,mod2,mod2b,mod3,mod4,mod5,mod6)
```

### Diagnostics
```{r}
# Show outputs
stargazer(mod1, mod2, mod3)

data.frame(FE_Region=rbind(AIC=AIC(mod1),BIC=BIC(mod1),
                           ResidVar=mod1@sigma^2),
           FE_Region2=rbind(AIC=AIC(mod1b),BIC=BIC(mod1b),
                            ResidVar=mod1b@sigma^2),
           FE_Govt=rbind(AIC=AIC(mod2),BIC=BIC(mod2),
                         ResidVar=mod2@sigma^2),
           FE_Govt2=rbind(AIC=AIC(mod2b),BIC=BIC(mod2b),
                          ResidVar=mod2b@sigma^2),
           FE_Govt_RE_Bank=rbind(AIC=AIC(mod3),BIC=BIC(mod3),
                                 ResidVar=mod3@sigma^2))
# they're all very similar so far, model 2b with fixed effects for government region
# and dependency ratio 2 has the lowest AIC and BIC, highest residual variance but
# they are all extremely comparable with this metric. Lower AIC and BIC is to be
# expected for mod2 and mod2b because they are estimating far fewer parameters
# ***CRITERIA INCREASED A LOT WHEN SPECIFYING WEIGHTS

AIC(mod4) # AIC is basically the same as the other models (which makes sense because of small variances)
AIC(mod5) # similar again, same explanation as above
AIC(mod6) # similar again, same explanation as above

# residual plots
par(mfrow=c(3,2))
qqnorm(resid(mod1),main="Model 1 Residuals")
qqline(resid(mod1)) 
plot(resid(mod1),fitted(mod1))
qqnorm(resid(mod2),main="Model 2 Residuals")
qqline(resid(mod2))
plot(resid(mod2),fitted(mod2))
qqnorm(resid(mod3),main="Model 3 Residuals")
qqline(resid(mod3))
plot(resid(mod3),fitted(mod3))

# check normality of random coefficients
par(mfrow=c(2,1))
qqnorm(ranef(model1)$HHIDPN[,1],main='Model1 Ints') # intercepts
qqline(ranef(model1)$HHIDPN[,1])
qqnorm(ranef(model2)$HHIDPN[,1],main='Model2 Ints')
qqline(ranef(model2)$HHIDPN[,1])
qqnorm(ranef(model3)$HHIDPN[,1],main='Model3 Ints')
qqline(ranef(model3)$HHIDPN[,1])
```

### EXTRA: EDA Based on Model Results
```{r}
# checking intuition that mara and geita are similar to dodoma based on mod1 and mod3 results
# they are all rural, which makes this fairly obvious
similar_regions <- subset(tanzania_1,tanzania_1$region %in% c("dodoma","geita","mara"))
similar_regions$region <- droplevels(similar_regions$region)

xtabs(~wealth_ind_q+region,data=similar_regions) # dodoma definitely poorer than the others...
xtabs(~hh_sex+region,data=similar_regions) # similar ratio of men to women
xtabs(~hh_educ+region,data=similar_regions) # fairly comparable education levels

similar_regions %>% # compute unweighted summary statistics
  group_by(region) %>%
  dplyr::summarize(
    mean_wealth = mean(wealth_ind),
    std_wealth = sd(wealth_ind),
    mean_members=mean(num_members),
    mean_head_age=mean(hh_age),
    mean_dep_ratio=mean(dep_ratio),
    mean_dep_ratio2=mean(dep_ratio2,na.rm=T)) %>% 
  arrange(desc(mean_wealth)) # dodoma with lower dep_ratio seems to be the main discrepancy

similar_regions %>% # compute weighted summary statistics
  group_by(region) %>%
  dplyr::summarize(
    mean_wealth = weighted.mean(wealth_ind,hh_weight),
    std_wealth = wtd.var(wealth_ind,hh_weight) %>% sqrt(),
    mean_members=weighted.mean(num_members,hh_weight),
    mean_head_age=weighted.mean(hh_age,hh_weight),
    mean_dep_ratio=weighted.mean(dep_ratio),
    mean_dep_ratio2=weighted.mean(dep_ratio2,na.rm=T)) %>% 
  arrange(desc(mean_wealth)) # dodoma has less members and (therefore?) lower dep_ratio

plot_grid(ggplot(similar_regions) + # very similar distributions of wealth
            geom_boxplot(mapping=aes(y=wealth_ind,group=region,fill=region)) +
            theme(axis.text.x = element_blank(),axis.ticks.x = element_blank(),
                  legend.position='none') +
            labs(title="Dispersion of Wealth Index across regions",y="Wealth Index"),
          ggplot(similar_regions) + 
            geom_density(mapping=aes(x=wealth_ind,group=region,fill=region),alpha=0.4) +
            labs(title="Dispersion of Wealth Index across regions",y="Wealth Index"))

rm(similar_regions)
```

