---
title: "707 Computational Examples"
author: "Alex Caramanico"
date: "`r Sys.Date()`"
output: pdf_document
---
# Load Libraries
```{r}
library(rstan)
library(R2jags)
library("bayesplot")
library("ggplot2")
library(loo)
library(cubature)
```
# 1
## Stan Code
```{r}
p1norm <- 'data {int y;}
parameters {real theta;}
model {
  y ~ normal(theta, 1);
  theta ~ normal(0, 10);}' # sampling function takes sigma, not sigma^2

p1cauch <- 'data {int y;}
parameters {real theta;}
model {
  y ~ normal(theta, 1);
  theta ~ cauchy(0, 1);}' 

p1stud <- 'data {int y;}
parameters {real theta;}
model {
  y ~ normal(theta, 1);
  theta ~ student_t(5,0,sqrt(5/3));}' # mu = 0 for v>1, sd = sqrt(v/(v-2)) for v>2
```
## Normal Prior
```{r}
p1norm.model <- stan_model(model_code=p1norm,model_name='Normal-Normal')
p1norm.draws <- sampling(p1norm.model,data = list(y = 10),
  chains = 2,cores = 2,iter = 10000,warmup = 5000,seed = 500)
```
### Display Results
```{r}
print(p1norm.draws)
p1norm.ex <- extract(p1norm.draws)
hist(p1norm.ex$theta,main='Normal-Normal Histogram',prob=TRUE)
lines(density(p1norm.ex$theta))
# High variance of prior leads to data dominating the prior, posterior mean is very close to observed data y=10
```
## Cauchy Prior
```{r}
p1cauch.model <- stan_model(model_code=p1cauch,model_name='Normal-Cauchy')
p1cauch.draws <- sampling(p1cauch.model,data = list(y = 10),
  chains = 2,cores = 2,iter = 10000,warmup = 5000, seed = 500)
```
### Display Results
```{r}
print(p1cauch.draws)
p1cauch.ex <- extract(p1cauch.draws)
hist(p1cauch.ex$theta,main='Normal-Cauchy Histogram',prob=TRUE)
lines(density(p1cauch.ex$theta))
```
## Student's T Prior
```{r}
p1stud.model <- stan_model(model_code=p1stud,model_name='Normal-Students')
p1stud.draws <- sampling(p1stud.model,data = list(y = 10),
  chains = 2,cores = 2,iter = 10000,warmup = 5000,seed = 500)
```
### Display Results
```{r}
print(p1stud.draws)
p1stud.ex <- extract(p1stud.draws)
hist(p1stud.ex$theta,main='Normal-Students Histogram',prob=TRUE)
lines(density(p1stud.ex$theta))
```

# 2 From Homework 1
# 3
## Data / Stan Code / Model
```{r}
set.seed(1726)
baseball_data <- read.csv("C:/Users/jacma/Downloads/baseball.csv")
early70 <- baseball_data$hat.P
careerBA <- baseball_data$x1
careerAB <- baseball_data$x2
realBA <- baseball_data$P
M <- length(early70)
avgBA <- sum(early70)/M
var <- avgBA*(1-avgBA)/45

p3code <- '
  data {
    int<lower=0> m;
    array[m] real<lower=0,upper=1> early70; 
    real<lower=0> D;
    matrix[m,3] xi; 
    }
  parameters {
    array[m] real<lower=0,upper=1> actualBA;
    real beta;
    real<lower=0> alpha;
    }
  model {
    for (i in 1:m){  
      early70[i] ~ normal(actualBA[i],sqrt(D));
      actualBA[i] ~ normal(xi[i,]*beta,sqrt(alpha)); 
      }
    beta ~ uniform(0,1); 
    alpha ~ inv_gamma(10,10); 
    }'
```

## a) For each player, compute the posterior mean and standard deviation.
***WARNING CONCERN***
```{r}
aux <- data.frame(1,baseball_data[,3],baseball_data[,4])
# ^ this is a matrix with 18 rows, 3 columns, that will be parsed for each player to incorporate the effect of the auxiliary variables
ex3.model <- stan_model(model_code=p3code,model_name='ex3')
ex3.draws <- sampling(ex3.model,
  data = list(m=M,early70=baseball_data[,2],xi=aux,D=var), 
  chains = 2,cores = 2,iter = 10000,warmup = 5000,seed = 567)
```
```{r}
print(ex3.draws)
```
## b) Using results in part a), obtain box plots of estimation errors 
$\hat{\theta_i}^{HB}- \theta_i$ and $\hat{\theta_i}- \theta_i$ in the same plot.
```{r}
post_means1 <- get_posterior_mean(ex3.draws)[1:18,3]
prediction_errs <- post_means1 - realBA
real_diff <- early70 - realBA
boxplot(prediction_errs,real_diff,ylab='Midseason      Model Predictions',
        xlab='Error',horizontal=T)
```
## c) Obtain the 95% credible intervals of true batting averages for all of the 18 players.
```{r}
# see part a), look at 2.5% and 97.5% columns 
```

# 4 *Models 6 and 7 are overestimating
For each of the following models, compute the summary measures ASD, ARSD, AAD, and ARAD, along with the error and relative error for each player. 
## **heavier tails might be better at capturing difference in player quality 
## (Clemente is not a <.270 hitter) 
### *** the priors are absolutely dominating the data in these models. I want it to distinguish between good players. don't want the expected result to bring them THAT close together. I could spend all day trying to fine tune this
## Summary Measures Function *take in the draws get_posterior_mean(DRAWS)[1:18,3]? 
```{r}
summary_measures <- function(Model,prediction,actualBA,n=M){
  ASD <- sum((prediction-actualBA)^2)/n
  ARSD <- sum((prediction-actualBA)^2/actualBA^2)/n
  AAD <- sum(abs(prediction-actualBA))/n
  ARAD <- sum(abs(prediction-actualBA)/actualBA)/n
  all_measures <- data.frame(Model,ASD,ARSD,AAD,ARAD)
  return(all_measures)
}

individual_errors <- function(prediction, actualBA){
  error <- prediction - actualBA
  relative_error <- (prediction-actualBA)/actualBA
  indie_errors <- data.frame(error,relative_error)
  return(indie_errors)
}
model0 <- summary_measures('Model 0 - Normal Approx to Binomial', post_means1,realBA,M)
#print(model0)
# print(individual_errors(post_means1,realBA))
```
## Model 1 Binomial-Beta Informative/Non-Informative
```{r}
# non-informative, uniform prior
m41codea <- '
  data {
    int<lower=0> n;
    int<lower=0> m;
    array[m] int<lower=0,upper=n> earlyhits;
    }
  parameters {
    array[m] real<lower=0,upper=1> actualBA;
    }
  model {
    for (i in 1:m){
      earlyhits[i] ~ binomial(n,actualBA[i]);
      actualBA[i] ~ uniform(0,1);
      }
    }'
# informative prior, beta(40,120) seems to be reasonable distribution of batting averages, large parameter values to make tails thinner
# hist(rbeta(10000,40,120))
m41codeb <- '
  data {
    int<lower=0> n;
    int<lower=0> m;
    array[m] int<lower=0,upper=n> earlyhits;
    }
  parameters {
    array[m] real<lower=0,upper=1> actualBA;
    }
  model {
    for (i in 1:m){
      earlyhits[i] ~ binomial(n,actualBA[i]);
      actualBA[i] ~ beta(40,120);
      }
    }'
# Jeffrey's prior is probably not reasonable because the majority of batting averages would not be near 0 or 1, which is the form of that prior (Beta(0.5,0.5)))
```
### Sampling and Display of Results
```{r}
earlyHits <- round(early70*45)
ex41.modela <- stan_model(model_code=m41codea,model_name='ex41a')
ex41.drawsa <- sampling(ex41.modela,
  data = list(n=45,m=M,earlyhits=earlyHits), 
  chains = 2,cores = 2,iter = 10000,warmup = 5000,seed = 567)
```
```{r}
ex41.modelb <- stan_model(model_code=m41codeb,model_name='ex41b')
ex41.drawsb <- sampling(ex41.modelb,
  data = list(n=45,m=M,earlyhits=earlyHits), 
  chains = 2,cores = 2,iter = 10000,warmup = 5000,seed = 567)
```
```{r}
# print(ex41.drawsa)
post_means2 <- get_posterior_mean(ex41.drawsa)[1:18,3]
prediction_errs2 <- post_means2 - realBA
model41a <- summary_measures('Model 1a - Binomial w/ Uniform Prior', post_means2,realBA,M)

# print(ex41.drawsb)
post_means3 <- get_posterior_mean(ex41.drawsb)[1:18,3]
prediction_errs3 <- post_means3 - realBA
model41b <- summary_measures('Model 1b - Binomial w/ Informative Beta Prior',
                             post_means3,realBA,M)
model1.summary <- rbind(model41a,model41b)

model41aI <- individual_errors(post_means2,realBA)
model41bI <- individual_errors(post_means3,realBA)
model1.summary
```

## Model 2 Binomial-Beta with Provided Priors
```{r}
m42codea <- '
  data {
    int<lower=0> n;
    int<lower=0> m;
    array[m] int<lower=0,upper=n> earlyhits;
    }
  parameters {
    array[m] real<lower=0,upper=1> actualBA;
    }
  model {
    for (i in 1:m){
      earlyhits[i] ~ binomial(n,actualBA[i]);
      actualBA[i] ~ beta(1,1);
      }
    }'
m42codeb <- '
  data {
    int<lower=0> n;
    int<lower=0> m;
    array[m] int<lower=0,upper=n> earlyhits;
    }
  parameters {
    array[m] real<lower=0,upper=1> actualBA;
    }
  model {
    for (i in 1:m){
      earlyhits[i] ~ binomial(n,actualBA[i]);
      actualBA[i] ~ beta(sqrt(n/4),sqrt(n/4));
      }
    }'
```
### Sampling and Display of Results
```{r}
ex42.modela <- stan_model(model_code=m42codea,model_name='ex42a')
ex42.drawsa <- sampling(ex42.modela,
  data = list(n=45,m=M,earlyhits=earlyHits), 
  chains = 2,cores = 2,iter = 10000,warmup = 5000,seed = 567)
```
```{r}
ex42.modelb <- stan_model(model_code=m42codeb,model_name='ex42b')
ex42.drawsb <- sampling(ex42.modelb,
  data = list(n=45,m=M,earlyhits=earlyHits), 
  chains = 2,cores = 2,iter = 10000,warmup = 5000,seed = 567)
```
```{r}
post_means4 <- get_posterior_mean(ex42.drawsa)[1:18,3]
prediction_errs4 <- post_means4 - realBA
model42a <- summary_measures('Model 2a - Binomial w/ Uniform Prior', post_means4,realBA,M)

post_means5 <- get_posterior_mean(ex42.drawsb)[1:18,3]
prediction_errs5 <- post_means5 - realBA
model42b <- summary_measures('Model 2b - Binomial w/ Beta(sqrt(n/4) Prior',
                             post_means5,realBA,M)

model2.summary <- rbind(model42a,model42b)

model41aI <- individual_errors(post_means2,realBA)
model41bI <- individual_errors(post_means3,realBA)
```

## Model 3 Binomial-Beta with Hyperpriors
```{r}
# I think higher values for the Beta parameters give a more desirable / "accurate" kernel on batting average, so Gamma hyperpriors also have large parameter values
m43code <- '
  data {
    int<lower=0> n;
    int<lower=0> m;
    array[m] int<lower=0,upper=n> earlyhits;
    }
  parameters {
    array[m] real<lower=0,upper=1> actualBA;
    real<lower=0> alpha;
    real<lower=0> beta;
    }
  model {
    alpha ~ gamma(50,5);
    beta ~ gamma(100,1);
    for (i in 1:m){
      earlyhits[i] ~ binomial(n,actualBA[i]);
      actualBA[i] ~ beta(alpha,beta);
      }
    }'
```
### Sampling and Display of Results
```{r}
ex43.model <- stan_model(model_code=m43code,model_name='ex43')
ex43.draws <- sampling(ex43.model,
  data = list(n=45,m=M,earlyhits=earlyHits), 
  chains = 2,cores = 2,iter = 10000,warmup = 5000,seed = 567)
```
```{r}
# print(ex43.draws)
post_means6 <- get_posterior_mean(ex43.draws)[1:18,3]
prediction_errs6 <- post_means6 - realBA
model3.summary <- summary_measures(
  'Model 3 - Binomial with Beta Prior, Gamma Hyperprior',post_means6,realBA,M)
```


## Model 4 Synthetic Model with arcsine Transformation
$Y_i=\sqrt{n}arcsin(2\hat{\theta_i}-1)$
$\eta_i=\sqrt{n}arcsin(2\hat{\theta_i}-1)$
$\hat{\theta_i}=\frac{1}{2}sin(\frac{Y_i}{\sqrt{n}}+1)$
```{r}
# non-informative, uniform prior
m44codea <- '
  data {
    int<lower=0> m;
    int<lower=0> n;
    array[m] real yi;
    }
  parameters {
    array[m] real<lower=-sqrt(n),upper=sqrt(n)> eta;
    real<lower=0> sigma;
    }
  model {
    for (i in 1:m){
      yi[i] ~ normal(eta[i],sigma);
      eta[i] ~ uniform(-sqrt(n),sqrt(n));
      }
    sigma ~ inv_gamma(10,10);
    }'
# .250 BA leads to yi of -3.5 --> mean..?
m44codeb <- '
  data {
    int<lower=0> m;
    int<lower=0> n;
    array[m] real yi;
    }
  parameters {
    array[m] real<lower=-sqrt(n),upper=sqrt(n)> eta;
    real<lower=0> sigma;
    }
  model {
    for (i in 1:m){
      yi[i] ~ normal(eta[i],sigma);
      eta[i] ~ normal(-3.5,0.5);
      }
    sigma ~ inv_gamma(100,100);
    }'
```
```{r}
transformed <- sqrt(45)*asin(2*baseball_data[,2]-1) # arcsin transformations
ex44.modela <- stan_model(model_code=m44codea,model_name='ex44a')
ex44.drawsa <- sampling(ex44.modela,
  data = list(m=M,yi=transformed,n=45), 
  chains = 2,cores = 2,iter = 10000,warmup = 5000,seed = 567)
```
```{r}
ex44.modelb <- stan_model(model_code=m44codeb,model_name='ex44b')
ex44.drawsb <- sampling(ex44.modelb,
  data = list(m=M,yi=transformed,n=45), 
  chains = 2,cores = 2,iter = 10000,warmup = 5000,seed = 567)
```
```{r}
# print(ex44.drawsa)
fittedyi <- get_posterior_mean(ex44.drawsa)[1:18,3]
m44fit <- 1/2*sin(fittedyi/sqrt(45)+1) # convert back to proportion
model44a <- summary_measures('Model 4a - Synthetic Model with Noninformative Prior', 
                             m44fit, realBA)

fittedyi2 <- get_posterior_mean(ex44.drawsb)[1:18,3]
m44fit2 <- 1/2*sin(fittedyi2/sqrt(45)+1) # convert back to proportion
model44b <- summary_measures('Model 4b - Synthetic Model with Informative Prior', 
                             m44fit2, realBA)
model4.summary <- rbind(model44a,model44b)
```

## Model 5 Regression Model
### Noninformative Priors 
```{r}
m45codea <- '
  data {
    int<lower=0> m;
    int<lower=0> n;
    array[m] real yi;
    array[m] real career;
  }
  parameters{
    real beta0;
    real beta1;
    real<lower=0> sigma;
  }
  transformed parameters{
    array[m] real thetai;
    for (i in 1:m) {thetai[i] = beta0 + beta1*career[i];}
  }
  model{
    beta0 ~ normal(0,100);
    beta1 ~ normal(0,100);
    for (i in 1:m){
      yi[i] ~ normal(thetai[i],sigma);
    }
  }
  '
```
```{r}
ex45.modela <- stan_model(model_code=m45codea,model_name='ex45a')
ex45.drawsa <- sampling(ex45.modela,
  data = list(yi=transformed,m=M,career=careerBA,n=45), 
  chains = 2,cores = 2,iter = 10000,warmup = 5000,seed = 567)
```
```{r}
# print(ex45.drawsa)
post_meansreg <- get_posterior_mean(ex45.drawsa)[4:21,3]
m45fit <- 1/2*sin(post_meansreg/sqrt(45)+1) # convert back to proportion
model45a <- summary_measures('Model 5a - Regression with Noninformative Priors',
                             m45fit,realBA,M)
```
### Informative Priors 
```{r}
# mean batting average, i.e. intercept, is informative Bayes
# slope according to early batting average would in theory be strictly positive, gamma, theta must remain between 0 and 1
m45codeb <- '
  data {
    int<lower=0> m;
    array[m] real y;
    array[m] real career;
  }
  parameters{
    real beta0;
    real beta1;
    real<lower=0> sigma;
  }
  transformed parameters{
    array[m] real thetai;
    for (i in 1:m) {thetai[i] = beta0 + beta1*career[i];}
  }
  model{
    beta0 ~ normal(-3.5,5);
    beta1 ~ normal(0,10);
    for (i in 1:m){y[i] ~ normal(thetai[i],sigma);}
  }'
```
```{r}
ex45.modelb <- stan_model(model_code=m45codeb,model_name='ex45b')
ex45.drawsb <- sampling(ex45.modelb,
  data = list(y=transformed,m=M,career=careerBA), 
  chains = 2,cores = 2,iter = 10000,warmup = 5000,seed = 567)
```
```{r}
# print(ex45.drawsb)
post_meansreg2 <- get_posterior_mean(ex45.drawsb)[4:21,3]
m45fit2 <- 1/2*sin(post_meansreg2/sqrt(45)+1) # convert back to proportion
model45b <- summary_measures('Model 5b - Regression with Informative Priors',
                             m45fit2,realBA,M)
model5.summary <- rbind(model45a,model45b)
```

## Model 6 Normal-Normal *predictions way too high
```{r}
m46codea <- '
  data {
    int<lower=0> m;
    array[m] real y;
    }
  parameters {
    array[m] real<lower=0> theta;
    real<lower=0,upper=1> mu;
    real<lower=0> tau;
  }
  model {
    for (i in 1:m){
      y[i] ~ normal(theta[i],1);
      theta[i] ~ normal(mu,tau);
      }
    mu ~ uniform(0,1);
    tau ~ uniform(0,100);
    }'
m46codeb <- '
  data {
    int<lower=0> m;
    array[m] real y;
    }
  parameters {
    array[m] real<lower=0> theta;
    real mu;
    real<lower=0> tau;
    }
  model {
    for (i in 1:m){
      y[i] ~ normal(theta[i],1);
      theta[i] ~ normal(mu,tau);
      }
    }'
m46codec <- '
  data {
    int<lower=0> m;
    array[m] real y;
    }
  parameters {
    array[m] real<lower=0> theta;
    real<lower=0> mu;
    real<lower=0> tau;
    }
  model {
    for (i in 1:m){
      y[i] ~ normal(theta[i],1);
      theta[i] ~ normal(mu,tau);
      }
    mu ~ inv_gamma(3,1);
    tau ~ inv_gamma(3,0.5);
    }'
```
```{r}
ex46.modela <- stan_model(model_code=m46codea,model_name='ex46a')
ex46.drawsa <- sampling(ex46.modela,
  data = list(m=M,y=transformed), 
  chains = 2,cores = 2,iter = 10000,warmup = 5000,seed = 567)
```
```{r}
ex46.modelb <- stan_model(model_code=m46codeb,model_name='ex46b')
ex46.drawsb <- sampling(ex46.modelb,
  data = list(m=M,y=transformed), 
  chains = 2,cores = 2,iter = 10000,warmup = 5000,seed = 567)
```
```{r}
ex46.modelc <- stan_model(model_code=m46codec,model_name='ex46c')
ex46.drawsc <- sampling(ex46.modelc,
  data = list(m=M,y=transformed), 
  chains = 2,cores = 2,iter = 10000,warmup = 5000,seed = 567)
```
```{r}
# print(ex46.drawsa)
model6postmeansa <- get_posterior_mean(ex46.drawsa)[1:18,3]
m46fit <- 1/2*sin(model6postmeansa/sqrt(45)+1) # convert back to proportion
model46a <- summary_measures('Model 6a - Normal/Normal with Noninformative Priors',
                             m46fit,realBA,M)
# print(ex46.drawsb)
model6postmeansb <- get_posterior_mean(ex46.drawsb)[1:18,3]
m46fit2 <- 1/2*sin(model6postmeansb/sqrt(45)+1) # convert back to proportion
model46b <- summary_measures('Model 6b - Normal/Normal with Jeffreys Priors',
                             m46fit2,realBA,M)

# print(ex46.drawsc)
model6postmeansc <- get_posterior_mean(ex46.drawsc)[1:18,3]
m46fit3 <- 1/2*sin(model6postmeansc/sqrt(45)+1) # convert back to proportion
model46c <- summary_measures('Model 6c - Normal/Normal with Inverse Gamma Priors',
                             m46fit3,realBA,M)
model6.summary <- rbind(model46a,model46b,model46c)
```

## Model 7 *predictions also too high
```{r}
m47codea <- '
  data {
    int<lower=0> m;
    array[m] real y;
    array[m] real<lower=0> career;
    }
  parameters {
    real beta0;
    real beta1;
    real<lower=0> tau;
    array[m] real<lower=0> theta;
  }
  transformed parameters{
    array[m] real eta;
    for (i in 1:m) {eta[i] = beta0 + beta1*career[i];}
  }  
  model {
    for (i in 1:m){
      y[i] ~ normal(theta[i],1);
      theta[i] ~ normal(eta[i],tau);
      }
    beta0 ~ uniform(-100,100);
    beta1 ~ uniform(-100,100);
    tau ~ inv_gamma(10,10);
    }'
m47codeb <- '
  data {
    int<lower=0> m;
    array[m] real y;
    array[m] real<lower=0> career;
    }
  parameters {
    real beta0;
    real beta1;
    real<lower=0> tau;
    array[m] real<lower=0> theta;
  }
  transformed parameters{
    array[m] real eta;
    for (i in 1:m) {eta[i] = beta0 + beta1*career[i];}
  }  
  model {
    for (i in 1:m){
      y[i] ~ normal(theta[i],1);
      theta[i] ~ normal(eta[i],tau);
      }
    }'
m47codec <- '
  data {
    int<lower=0> m;
    array[m] real y;
    array[m] real<lower=0> career;
    }
  parameters {
    real beta0;
    real beta1;
    real<lower=0> tau;
    array[m] real<lower=0> theta;
  }
  transformed parameters{
    array[m] real eta;
    for (i in 1:m) {eta[i] = beta0 + beta1*career[i];}
  }  
  model {
    for (i in 1:m){
      y[i] ~ normal(theta[i],1);
      theta[i] ~ normal(eta[i],tau);
      }
    beta0 ~ inv_gamma(10,10);
    beta1 ~ inv_gamma(10,10);
    tau ~ inv_gamma(10,10);
    }'
```
```{r}
ex47.modela <- stan_model(model_code=m47codea,model_name='ex47a')
ex47.drawsa <- sampling(ex47.modela,
  data = list(m=M,y=transformed,career=careerBA), 
  chains = 2,cores = 2,iter = 10000,warmup = 5000,seed = 567)
```
```{r}
ex47.modelb <- stan_model(model_code=m47codeb,model_name='ex47b')
ex47.drawsb <- sampling(ex47.modelb,
  data = list(m=M,y=transformed,career=careerBA), 
  chains = 2,cores = 2,iter = 10000,warmup = 5000,seed = 567)
```
```{r}
ex47.modelc <- stan_model(model_code=m47codec,model_name='ex47c')
ex47.drawsc <- sampling(ex47.modelc,
  data = list(m=M,y=transformed,career=careerBA), 
  chains = 2,cores = 2,iter = 10000,warmup = 5000,seed = 567)
```

```{r}
m7postmeansa <- get_posterior_mean(ex47.drawsa)[4:21,3]
m47fit <- 1/2*sin(m7postmeansa/sqrt(45)+1) # convert back to proportion
model7a <- summary_measures('Model 7a - Normal/Normal Regression with Uniform Priors',
                            m47fit, realBA)
m7postmeansb <- get_posterior_mean(ex47.drawsb)[4:21,3]
m47fit2 <- 1/2*sin(m7postmeansb/sqrt(45)+1) # convert back to proportion
model7b <- summary_measures('Model 7b - Normal/Normal Regression with Jeffreys Priors',
                            m47fit2, realBA)

m7postmeansc <- get_posterior_mean(ex47.drawsc)[4:21,3]
m47fit3 <- 1/2*sin(m7postmeansc/sqrt(45)+1) # convert back to proportion
model7c <- summary_measures('Model 7c - Normal/Normal Regression with IG Priors',
                            m47fit3, realBA)

model7.summary <- rbind(model7a,model7b,model7c)
cbind(m47fit,m47fit2,m47fit3)
```

## Overall Results
```{r}
rbind(model1.summary,model2.summary,model3.summary,model4.summary,
      model5.summary,model6.summary,model7.summary)
# posterior_cp <- as.array(ex45.drawsb)
# np_cp <- nuts_params(ex45.drawsb)
# color_scheme_set("darkgray")
# mcmc_parcoord(posterior_cp, np = np_cp)
```
### For the top three models, provide detailed analysis. The output should have different diagnostic plots you think are important, plot of some sample density (say, θi for the three players I mentioned earlier),regression coefficients, variance components, posterior means, posterior variances, credible intervals, etc Models 4-7 need to be checked, especiall 5-7 in their implementation
# 5
## a)
```{r}
library(ggplot2)
prior1 <- data.frame(value = rbeta(10000,1,1))
prior2 <- data.frame(value = rbeta(10000,0.5,0.5))
prior3 <- data.frame(value = rbeta(10000,0,0))
prior1$prior <- 'Uniform'
prior2$prior <- 'Jeffreys'
prior3$prior <- 'Improper'
prior_dists <- rbind(prior1,prior2,prior3)
ggplot(data=prior_dists, aes(x=value, fill = prior)) + 
  geom_density(alpha = 0.2) + 
  ggtitle('Prior Distributions')
```
## b) and c)
```{r}
# Binomial Posterior with Beta prior is Beta(alpha + y, beta + n - y)
N <- 493472
Y <- 241945
post1 <- data.frame(value = rbeta(10000,1+Y,1+N-Y)) 
post2 <- data.frame(value = rbeta(10000,0.5+Y,0.5+N-Y))
post3 <- data.frame(value = rbeta(10000,0+Y,0+N-Y))
post1$prior <- 'Uniform'
post2$prior <- 'Jeffreys'
post3$prior <- 'Improper'
posterior_dists <- rbind(post1,post2,post3)
ggplot(posterior_dists, aes(value, fill = prior)) + 
  geom_density(alpha = 0.2) + 
  ggtitle('Posterior Distributions')
```
## d) and e)
```{r}
post_jeffrey <- rbeta(2000,0.5+Y,0.5+N-Y)
hist(post_jeffrey)
lines(density(post_jeffrey))
# e)
mean(post_jeffrey) # i. E[theta|y=241945]
var(post_jeffrey) # ii. V[theta|y=241945]
pbeta(0.5,0.5+Y,0.5+N-Y,lower.tail=FALSE) # iii. P(theta>0.5|y=241945]
jeff_cred <- qbeta(c(0.975,.025),0.5+Y,0.5+N-Y,
                   lower.tail=FALSE) # iv. 95% credible interval
jeff_cred
```
## f)
```{r}
p5code <- 'data {int<lower = 0> y;
                int<lower = 0> n;} 
    parameters {real theta;}
    model {y~binomial(n,theta);
           theta~beta(0.5,0.5);}' 
p5.model <- stan_model(model_code=p5code,model_name='Jeffrey Beta')
p5.draws <- sampling(
  p5.model,
  data = list(y = Y,n=N),
  chains = 2,
  cores = 2,
  iter = 10000,
  warmup = 5000,
  seed = 500
)
```
```{r}
p5results <- extract(p5.draws)
pfem <- p5results$theta
credint <- quantile(pfem,c(0.025,0.975))
cbind(Mean=mean(pfem),Variance=var(pfem),
      P=sum(pfem>0.5)/length(pfem),CredLow=credint[1],CredUp=credint[2])
# The sample size is so large that it is absolutely dominating the prior
```

# 6 * MISSING EFFECTIVE MODEL SIZE AND DIC
## a)
```{r}
patient <- 1:12
x6 <- c(102.4,103.2,101.9,103,101.2,100.7,
      102.5,103.1,102.8,102.3,101.9,101.4) # initial values
y6 <- c(99.6,100.1,100.2,101.1,99.8,100.3,
      101,100.1,100.7,101.1,101.3,100.2) # after values
p6data <- as.data.frame(cbind(patient,x6,y6))
p6data$z6 <- x6-y6 # difference, or reduction
ggplot(data=p6data,mapping=aes(x=x6,y=z6))+
  geom_point()+
  geom_smooth(method='lm')+
  xlab('Initial Temperature')+
  ylab('Temperature Reduction') +
  ggtitle('Reduction in temperature vs. initial temperature')
# general increasing trend, linear model may be appropriate. implies higher initial temperatures lead to greater reductions after aspirin is given
```
## b) Fit SLR Model 
$z_i = \beta_0 + \beta_1*(x_i-\bar{x}) + \epsilon_i, \epsilon \sim N(0,\sigma^2)$
```{r}
scaledx <- p6data$x6-mean(p6data$x6)
slr.stan.code <- '
  data{
    int<lower = 0> n;
    array[n] real x;
    array[n] real z;
    }
  parameters{
    real beta0;
    real beta1;
    real<lower=0> tau;
    }
  model{
    for (i in 1:n){
      z[i] ~ normal(beta0 + beta1*x[i],tau);
      }
    }
  generated quantities{
    array[n] real log_lik;
    array[n] real mu;
    for (i in 1:n){
      mu[i] = beta0 + beta1*x[i];
      log_lik[i] = normal_lpdf(z[i]|mu[i],tau);
      }
    }'
ex6.model = stan_model(model_code=slr.stan.code,model_name='ex6')
ex6.draws = sampling(ex6.model,data = list(n=12,x=scaledx,z=p6data$z6),
  chains = 2,cores = 2,iter = 10000,warmup = 5000,seed = 303)
```
### Results, Effective Model Size and DIC
$\bar{D}=mean(-2*log(P(Y|theta))$
```{r}
print(ex6.draws) 
# 95% credible interval for initial temperature effect: (0.35,1.35), results that match the intuition of the researchers. This indicates that higher initial values lead to a greater reduction in temperature.
log_lik1 <- extract_log_lik(ex6.draws, merge_chains = FALSE)
mean(-2*(get_posterior_mean(ex6.draws)[4:15]))
summary(ex6.draws)
z6 <- x6-y6
-2*prod(dlnorm(z6))
```
## c) New patient arrives with initial temperature of 100. Draw predictive density of
$Z_{13}$, predicted reduction in temperature for the patient, and point estimate for $P(Z_{13}>0|Z_{1},...,Z_{13})$, the probability their temperature is reduced
```{r}
ex6_predict <- '
  data{
    int<lower=0> n;
    real xNew;
    array[n] real x;
    array[n] real z;
    }
  parameters{
    real beta0;
    real beta1;
    real<lower=0> tau;
    }
  model{
    for (i in 1:n){
      z[i] ~ normal(beta0 + beta1*x[i],tau);
      }
    }
  generated quantities{
    real z13;
    z13 <- beta0 + beta1*(xNew);
  }
'
ex6pred.model <- stan_model(model_code=ex6_predict,model_name='ex6pred')
ex6pred.draws <- sampling(ex6pred.model,
                         data = list(n=12,xNew=100-mean(p6data$x6),x=scaledx,z=p6data$z6),
  chains = 2,cores = 2,iter = 10000,warmup = 5000,seed = 303)
```
$z_{new}=1.74+0.86(x_{new}-\bar{x})$
```{r}
print(ex6pred.draws)
prediction.ex <- extract(ex6pred.draws)
newzedz <- prediction.ex$z13
ggplot(data.frame(newzedz), aes(x = newzedz)) + 
  geom_density(fill = "blue", color="black", alpha = 0.2) +
  ggtitle('Posterior Predictive Density for Xnew=100') +
  xlab('Znew prediction') +
  geom_vline(xintercept=0,linetype='dashed',color='red') +
  annotate('text',x=0.75,y=0.4,label=sum(newzedz>0)/length(newzedz))
```

## d) Add quadratic term to model
$\beta_2(x_i-\bar{x})^2$
```{r}
slr.stan.code2 <- '
  data{
    int<lower = 0> n;
    array[n] real x;
    array[n] real z;
    real xNew;
    }
  parameters{
    real beta0;
    real beta1;
    real beta2;
    real<lower=0> tau;
    }
  model{
    for (i in 1:n){
      z[i] ~ normal(beta0 + beta1*x[i] + beta2*x[i]^2,tau);
      }
    }
  generated quantities{
    real z13;
    z13 <- beta0 + beta1*(xNew);
    array[n] real log_lik;
    array[n] real mu;
    for (i in 1:n){
      mu[i] = beta0 + beta1*x[i]+beta2*x[i]^2;
      log_lik[i] = normal_lpdf(z[i]|mu[i],tau);
      }
    }
'
ex6.model2 <- stan_model(model_code=slr.stan.code2,model_name='ex6b')
ex6.draws2 <- sampling(ex6.model2,data = list(n=12,x=scaledx,z=p6data$z6,xNew=-2.2),
  chains = 2,cores = 2,iter = 10000,warmup = 5000,seed = 303)
```
### Recheck effective model size and DIC. Apply to new data from part c) as well. 
```{r}
print(ex6.draws2)
# The quadratic term does not appear to be necessary, the 95% credible interval includes zero suggesting it does not have a significant effect on the reduction in temperature. Looking at the data, there does not appear to be much of a curvilinear trend, which supports this result. 
prediction.ex2 <- extract(ex6.draws2)
beta2.ex <- prediction.ex2$beta2
ggplot(data.frame(beta2.ex),aes(x=beta2.ex)) + 
  geom_density(fill='orange',color='black',alpha=0.2) +
  ggtitle('Posterior Distribution of Quadratic Parameter') +
  xlab('beta2') +
  geom_vline(xintercept=0,linetype='dashed',color='red') +
  annotate('text',x=-.5,y=0.4,label=sum(beta2.ex<0)/length(beta2.ex))

newzedz2 <- prediction.ex2$z13
ggplot(data.frame(newzedz), aes(x = newzedz)) + 
  geom_density(fill = "blue", color="black", alpha = 0.2) +
  ggtitle('Posterior Predictive Density for Xnew=100 (Quadratic Term)') +
  xlab('Znew prediction') +
  geom_vline(xintercept=0,linetype='dashed',color='red') +
  annotate('text',x=0.75,y=0.4,label=sum(newzedz2>0)/length(newzedz2))
```
# 7
The posterior pdf is given by
$p(\theta|y)=\frac{\prod_{i=1}^5[1+(\theta-y_i)^2]^{-1}I(\theta>0)}{\int_0^\infty\prod_{i=1}^5[1+(\theta-y_i)^2]^{-1}d\theta}$
## a) Numerical Integration
```{r}
y7 <- c(4,5.5,7.5,4.5,3)
posterior7 <- function(theta,y) {prod(1/(1+(theta-y)^2))*ifelse(theta>0,1,0)}

integral <- adaptIntegrate(function(theta) posterior7(theta, y7),
                           lowerLimit = 0,upperLimit = Inf)
normconst <- integral$integral

post_normed <- function(theta, y) {posterior7(theta, y7) / normconst}

postmeanNI <- adaptIntegrate(function(theta)
  theta * post_normed(theta, y7), lowerLimit = 0, upperLimit = Inf)$integral

postvarNI <- adaptIntegrate(function(theta)
  (theta - postmeanNI)^2 * post_normed(theta, y7),
  lowerLimit = 0,upperLimit = Inf)$integral
NIresults<- data.frame(Approach='Numerical Integration',
                       PosteriorMean=postmeanNI,PosteriorVariance=postvarNI)
```
## b) Monte Carlo
```{r}
thetas <- runif(10000, 0, 100)
# calculate posterior for each sample
post_samples <- sapply(thetas, function(theta) posterior7(theta, y7))

# normalize the posterior samples
post_sampNormed <- post_samples / sum(post_samples)

postmeanMC <- sum(thetas * post_sampNormed)
postvarMC <- sum((thetas - postmeanMC)^2 * post_sampNormed)
MCresults<- data.frame(Approach='Monte Carlo',
                       PosteriorMean=postmeanMC,PosteriorVariance=postvarMC)
```
## c) Importance Sampling
```{r}
ISthetas <- rnorm(10000,mean(y7),sd(y7))
imp_weights <- sapply(ISthetas, function(theta) posterior7(theta,y7) / 
                        dnorm(theta,mean(y7),sd(y7)))
imp_weights_normalize <- imp_weights / sum(imp_weights)
postmeanIS <- sum(ISthetas*imp_weights_normalize)
postvarIS <- sum((ISthetas-postmeanIS)^2*imp_weights_normalize)
ISresults<- data.frame(Approach='Importance Sampling',
                       PosteriorMean=postmeanIS,PosteriorVariance=postvarIS)
```
## d) Rejection Sampling
```{r}
RStheta <- rnorm(10000,mean(y7),sd(y7))
Paccept <- sapply(RStheta, function(theta) posterior7(theta,y7) /
                    (50*dnorm(theta,mean(y7),sd(y7)))
                  )
acc_samps <- RStheta[runif(10000)<Paccept]
RSresults<- data.frame(Approach='Rejection Sampling',
                       PosteriorMean=mean(acc_samps),PosteriorVariance=var(acc_samps))
```
## e) Markov Carlo Monte Chain
```{r}
m7code <- '
  data{
    int<lower=0> n;
    array[n] real y;
  }
  parameters{
    real theta;
  }
  model{
    for (i in 1:n){
      y[i]~cauchy(theta,1);
    }
  }
'
p7model <- stan_model(model_code=m7code,model_name='Problem7 MCMC')
p7draws <- sampling(p7model, data=list(y=y7,n=5),
                    chains=2,cores=2,iter=10000,warmup=5000,seed=901)
```
```{r}
MCMCmean <- get_posterior_mean(p7draws)[1,3]
MCMCvar <- var(extract(p7draws)$theta)
MCMCresults<- data.frame(Approach='MCMC',
                       PosteriorMean=MCMCmean,PosteriorVariance=MCMCvar)
```
## All results
```{r}
rbind(NIresults,MCresults,ISresults,RSresults,MCMCresults)
```

