---
title: "705 Homework 3"
author: "Alex Caramanico"
date: "`r Sys.Date()`"
output: pdf_document
---
# 1
```{r}
EucDist <- function(x=0,y=0){
  # i. Check vectors have same dimensionality
  if (length(x)==length(y)){return(sqrt(sum((x-y)^2)))}
  # ii. If mismatched dimensionality
  else{print('Vectors need to be in the same dimension!')}
}

EucDist(x=rnorm(3,0,1),y=rnorm(5,0,1))
N <- 1000
distance_est <- rep(0,N)
for (i in 1:N){
  distance_est[i] <- EucDist(x=rnorm(4,0,5),y=rnorm(4,0,5))
}
mean(distance_est)
```
I did 1000 replications of the function to come up with a LLN estimate for the
expected difference.

# 2-3
```{r warning=FALSE}
framework <- function(data_object,percent_missing=0.2){
  # i. Converts any column named ”Month” from NUMERIC into FACTOR type
  data_object$Month <- as.factor(data_object$Month)
  
  # ii. Converts any column named ”Day” from NUMERIC to CHARACTER type
  data_object$Day <- as.character(data_object$Day)

  # iii. Removes any column with more than 20% missing entries
  # subset df by negating logical test: 1 if total missing exceeds threshold
  num_obs <- nrow(data_object)
  threshold <- num_obs*percent_missing
  na_indices <- is.na(data_object[1:num_obs,])
  total_missing <- apply(na_indices,2,sum)
  data_object <- (data_object[,-(total_missing>threshold)])
  
  # iv. Replaces any remaining missing entries in NUMERIC columns with mean 
  val_columns <- unlist(lapply(data_object, is.numeric), use.names = FALSE) 
  numerics <- data_object[,val_columns]
  NA_index <- which(is.na(numerics), arr.ind = TRUE)
  # colMeans computes means, removing NA. impute mean acc to column place
  numerics[NA_index] <- colMeans(numerics, na.rm = TRUE)[NA_index[,"col"]]
  data_object[,val_columns] <- numerics # update original df
  
  # v. Produces a matrix called DataSum containing quartiles and mean
  DataSum <- matrix(apply(numerics,2,summary),ncol=ncol(numerics))
  
  
  # list: adjusted df, Datasum, two-column matrix with coords of imputed values
  items <- list(data_object,DataSum,NA_index)
  return(items)
}
start <- Sys.time()
data <- framework(airquality)
Sys.time() - start
data
```

# 4-8 https://r4ds.hadley.nz/data-visualize#exercises
## Section 1.2.5 exercises
### 1. How many rows are in penguins? How many columns?
```{r}
library(ggplot2)
library(palmerpenguins)
data(penguins)
dim(penguins)
```
The dim() function tells us the penguins data frame is 344 rows by 8 columns.

### 3. Scatterplot bill_depth_mm (y) vs. bill_length_mm (x). Describe relationship
```{r}
ggplot(data=penguins) + geom_point(mapping=aes(x=bill_length_mm,y=bill_depth_mm))
```
The relationship between the variables is somewhat difficult to ascertain with
the above plot, although there appears to be a *slight* positive relationship 
between bill length and bill depth. That said, there appear to be two separate 
groups, in that if we were to add lines of best fit they would both have 
positive slopes but different intercepts. 

### 4. Scatterplot of species vs. bill_depth_mm? What should geom be?
```{r}
ggplot(data=penguins) + geom_point(mapping=aes(x=species,y=bill_depth_mm))
ggplot(data=penguins) + geom_boxplot(mapping=aes(x=bill_depth_mm,fill=species)) 
ggplot(data=penguins) + geom_density(aes(x=bill_depth_mm,fill=species),alpha=0.2)
```
Scatterplots require the use of geom_point(), the plot groups the penguins by 
species and plots the bill depth on the y axis, which could be useful to infer
the spread and location of the variable for each of the species. geom_boxplot
may be a better choice, as demonstrated above, because it more intuitively shows
the dispersion of values for each group. geom_density may be even more useful 
because it allows the user to visualize any overlap in the distributions of the
variables.

### 5. Why does the following give an error and how would you fix it?
```{r}
# ggplot(data = penguins) + geom_point()
```
This gives an error because nothing is mapped to the x and y coordinates, so
geom_point() does not know what to do. The way to fix this would be to add 
mapping=aes(x= something, y=something_else), with the somethings being variables
of interest.

### 6. What does na.rm do in geom_point()? Its default value is? Plot when = TRUE
```{r}
ggplot(data=penguins) + 
  geom_point(mapping=aes(x=species,y=bill_depth_mm,color=species),na.rm=TRUE)
```
The na.rm parameter simply removes any of the "na" or missing values from the 
data set and plots the rest. The default value is FALSE, so that a warning is 
given to the user that the data has some missing values. When the parameter is 
true, the missing values are still removed but a warning is not given.

### 7. Add caption to previous plot “Data come from the palmerpenguins package.”
```{r}
ggplot(data=penguins) + 
  geom_point(mapping=aes(x=species,y=bill_depth_mm,color=species),na.rm=TRUE) +
  labs(caption='Data come from the palmerpenguins package.')
```

### 8. Recreate the following visualization. What aesthetic should bill_depth_mm be mapped to? And should it be mapped at the global level or at the geom level?
```{r}
ggplot(data=penguins) + 
  geom_point(mapping=aes(x=flipper_length_mm,y=body_mass_g,color=bill_depth_mm),na.rm=T) +
  geom_smooth(mapping=aes(x=flipper_length_mm,y=body_mass_g),na.rm=T)
```
bill_depth_mm needs to be mapped to the color aesthetic within geom_point to
recreate the above figure. It should be mapped at the geom level rather than the
global level so that the points themselves are actually colored.

## Section 1.4.3 
### 1. Bar plot of penguin species (species on y). How is this plot different?
```{r}
ggplot(data=penguins) + geom_bar(mapping=aes(y=species))
```
Assigning species to the y-axis simply flips the plot from what one may expect.
That is, species with more members in the data set have bars reaching further to
the right of the plot, rather than further upward. Tt can be seen that the most 
populous species in the data set is Adelie, and the least populous is Chinstrap.

### 2. How are these two plots different? Which aesthetic is more useful?
```{r}
ggplot(penguins, aes(x = species)) +
  geom_bar(color = "red")
ggplot(penguins, aes(x = species)) +
  geom_bar(fill = "red")
```
color simply changes the outline  of the bars, whereas fill changes the entire
bar into a different color. Either may be more useful depending on context, as 
you may want to outline the bars according to some factor and fill them 
according to another. Generally, the fill aesthetic seems more useful for 
actually changing the color of the bars themselves.

### 3. What does the bins argument in geom_histogram() do?
```{r}
ggplot(data=penguins) + geom_histogram(mapping=aes(x=bill_depth_mm,fill=species),bins=5,alpha=0.5)
ggplot(data=penguins) + geom_histogram(mapping=aes(x=bill_depth_mm,fill=species),bins=30,alpha=0.5) 
```
Bins effectively determines the number of bars in the histogram. If there are
10 bins, the data will be sorted into 10 groups, or bins, which corresponds to
10 bars in the plot. The number of bins does not always correspond to the number
of bars exactly, but it is a good indicator of how far apart the data must be to
be grouped separately.

### 4. Make histogram of carat variable. What binwidth is the most interesting?
```{r,warning=FALSE,message=FALSE}
library(tidyverse)
data(diamonds)
ggplot(data=diamonds) + geom_histogram(mapping=aes(x=carat)) +
  ggtitle('Carat Histogram with 30 bins (default)')
ggplot(data=diamonds) + geom_histogram(mapping=aes(x=carat),bins=5) +
  ggtitle('Carat Histogram with 5 bins')
ggplot(data=diamonds) + geom_histogram(mapping=aes(x=carat),bins=50) +
  ggtitle('Carat Histogram with 50 bins')
ggplot(data=diamonds) + geom_histogram(mapping=aes(x=carat),bins=100) +
  ggtitle('Carat Histogram with 100 bins')
ggplot(data=diamonds) + geom_histogram(mapping=aes(x=carat),bins=1000) +
  ggtitle('Carat Histogram with 1000 bins')
```
We can immediately rule out using a very low number of bins (e.g. 5) because
that makes data inference nearly impossible. Depending upon the level of detail
desired, the "most interesting" results can be seen with number of bins greater
than 30. The more bins, the more "spikes" in the data are revealed, which 
suggests the data is possibly multimodal. The last plot also reveals interesting
behavior (despite probably being a little *too* fine-detailed), in that there
are repeated values with high density, followed by a repeated exponential decay
pattern. For instance, one might see this behavior between carat values 0.5-0.7,
then again from 0.7 to 0.9. These values are approximate, but this pattern is
as if multiple $c^{-x+k}, c>0$ with $c,k\in\mathbb{R}^1$ functions are being
combined side-by-side.