---
title: "Stat 705 HW 4"
author: "Alex Caramanico"
date: "`r Sys.Date()`"
output: pdf_document
---
This document is very long because I have tried to go into great detail with 
discussions and analysis. \

# Problem 2.1 \
Cauchy PDF: $\frac{1}{s\pi(1+(x-\theta)/s)^2}$ $s=1$ \
Density: $f(x)=\frac{1}{\pi(1+(x-\theta)^2)}$\
Likelihood: $L(\theta)=\prod_{i=1}^{n}\frac{1}{\pi(1+(x-\theta)^2)}$\
Log-likelihood: $l(\theta)=-nlog(\pi)-\sum_{i=1}^{n}log(1+(x-\theta)^2)$\
Log-likelihood with our data (for visualization purposes):\
$-20log(\pi) - (log(1+(1.77-\theta)^2)+...+log(1+(43.21-\theta)^2)$
```{r,warning=FALSE}
library(ggplot2)
options(scipen=999)

# From lecture notes
Gradmat<-function(x, f, eps = 1e-06){
  # Calculate the difference-quotient approx gradient (matrix) of an arbitrary 
  # input (vector) function "f" at "x"
  dd = length(x)
  aa = length(f(x))
  epsmat = (diag(dd) * eps)/2
  gmat = array(0, dim = c(aa, dd))
  for(i in 1:dd){
    gmat[, i] <- (f(x + epsmat[, i]) -
    f(x - epsmat[, i]))/eps}
  if(aa > 1) gmat else c(gmat)
}

Newt_Raph_MLE<-function(guess,f, max.iter=25, tol = 1e-05, eps=1e-06){
  gradient <- function(x) Gradmat(x, f, eps)
  hessian <- function(x) Gradmat(x, gradient, eps)
  old <- guess
  new <- old - solve(hessian(old), gradient(old))
  it <- 1
  while(it < max.iter & sqrt(sum((new - old)^2)) >tol){
    old <- new
    new <- old - solve(hessian(old), gradient(old))
    it <- it + 1
  }
    list(Steps = it, Initial = guess, Estimate = new,fval = f(new))
}
```

## a. Newton-Raphson 
```{r}
# observed values
values <- c(1.77,-0.23,2.76,3.8,3.47,56.75,-1.34,4.24,-2.44,3.29,
       3.71,-2.4,4.53,-0.07,-1.05,-13.87,-2.53,-1.75,0.27,43.21)
# Plot Log-likelihood
theta <- seq(-10,10,by=.1)
log_likelihood <- function(theta) sum(dcauchy(values, location=theta,scale=1,log=TRUE))
# analytical first derivative of log likelihood
log_deriv <- function(theta) 2*sum((values-theta)/(1+(values-theta)^2))

# call above function with theta and observed values
plot(theta, unlist(lapply(theta, log_likelihood)),
     xlab='theta',ylab='log-likelihood', main='Log-likelihood of Cauchy(theta,1)')
```

```{r}
# the mean is about 5.1
# -11 and 38 lead to errors, "system is exactly singular" --> very poor guesses
test_values <- c(-1,0,1.5,4,4.7,7,8,mean(values))
results_df <- data.frame()
# test all values via for loop, index functional output to set up columns
# values updated according to ith row
for (i in 1:length(test_values)){
  newt_raph_results <- Newt_Raph_MLE(test_values[i],log_likelihood)
  results_df[i,1] <- newt_raph_results[1]
  results_df[i,2] <- newt_raph_results[2]
  results_df[i,3] <- newt_raph_results[3]
  results_df[i,4] <- log_deriv(as.numeric(newt_raph_results[3]))
}
results_df

# checking code
optimize(function(theta) -sum(dcauchy(values, location=theta, log=TRUE)),  c(-100,100)) 
```
The widely-varying results were to be expected based on the range of initial
guesses. -11 and 38 cause an error and were thus excluded from the code, which
suggests these initial values take a bad initial step and quickly diverge. The 
specific error is "system is exactly singular".\
The most common estimate among these guesses was -0.1922866, which matches the
result from the optimize() function. There are clearly a number of local maxima
for the objective function, which leads to several possible results for the 
estimate of $\theta$. The mean proves to be a solid estimate, it converges in 8
steps. Interestingly, it reaches a unique local maxima. It should also be noted
that the objective function evaluated at estimate of -0.1922866 gives the 
closest values to 0. Initial guesses closer to the actual maxima converge in
less steps, which is why values 7, 8, and even the mean, converge to a different
optimum. \

## b. Bisection Method
```{r}
bisect <- function(bounds,fun,eps=1e-05,max.iter=50){
  # bounds should be a vector, done this way to facilitate testing
  # fun is the function to optimize
  lower <- bounds[1]
  upper <- bounds[2]
  initial_up <- upper
  initial_low <- lower
  theta_guess <- (upper+lower) / 2
  it <- 1
  while (it < max.iter & abs(fun(theta_guess))> eps) {
    # bisection algorithm 
    if (fun(lower)*fun(theta_guess)<=0){upper <- theta_guess}
    else{lower<-theta_guess}
    theta_guess <- (upper + lower) / 2
    it <- it + 1
  }  
  list(Estimate=as.numeric(theta_guess), Steps=it, InitialLow=initial_low,InitialUp=initial_up)
}
bisection_test <- list(c(-1,1),c(0,2),c(-4,4),c(-40,40),c(1,1.2))
sapply(bisection_test,bisect,fun=log_deriv)
```
Overall, the bisection method takes much more steps than the Newton-Raphson 
algorithm (provided the latter has a solid initial guess). Of course, the wider 
the interval, the more steps we expect the algorithm to take. On the other hand,
a bad narrow interval (as demonstrated in the fifth row) can also lead to poor
performance. The estimates of -0.192 and 1.713 here match those obtained via 
Newton-Raphson. \

## c. Fixed-point iteration
```{r}
vectorized_loglike <- Vectorize(log_likelihood,'theta')

fixed_point <- function(guess, f, alpha=1, max.iter=25, eps=1e-05){
  # new theta is simply G(x) evaluated at old theta
  # If G(x)=g'(x)+x ---> |g''(x)+1|<1
  # G(x) = alpha*g'(x) + x
  # choose alpha so that |alpha*g''(x)+1|<1 on [a,b] 
  xi <- guess
  i <- 0
  gradient <- function(guess){return(Gradmat(guess,f,eps))}
  hessian <- function(guess) Gradmat(guess,gradient,eps)
  while (i < max.iter & abs(1+alpha*hessian(xi)) < 1){
    xi <- xi + alpha*gradient(xi)
    i <- i+1
  }  
  list(Initial=guess,Alpha=alpha,Estimate=xi,Steps=i)
}
alphas <- c(1,0.64,0.25,0.01)
guess1 <- sapply(alphas,fixed_point,guess=-1,f=vectorized_loglike)
guess2 <- sapply(alphas,fixed_point,guess=0,f=vectorized_loglike)
guess3 <- sapply(alphas,fixed_point,guess=4,f=vectorized_loglike)
cbind(guess1,guess2,guess3)
```
The choice of alpha is very important, it appears that alpha of 0.25 is the most
appropriate choice as it converges to the solution of -.1922866 as obtained by
previous methods. The maximum number of steps is being reached in arriving at 
this solution, regardless of the initial guess (at least for -1 and 0). Much
more analysis, including varied initial guesses and perhaps a range of alphas,
can be done to ascertain the performance of the fixed point method.\

## d. Secant Method
```{r}
secant <- function(initial, f, max.iter=500, tol=1e-05){
  theta_0 <- as.numeric(initial[1])
  theta_1 <- as.numeric(initial[2])
  for (i in 1:max.iter){
    theta_2 <- theta_1 - f(theta_1) / ((f(theta_1)-f(theta_0)) / (theta_1-theta_0)) 
    if (abs(theta_2-theta_1) < tol) {
      return(list(Estimate=theta_2,Func_Val=log_deriv(theta_2),Initial_1=initial[1],Initial_2=initial[2]))}
    theta_0 <- theta_1
    theta_1 <- theta_2
  }
}
secant_test <- list(c(-2,1),c(-3,3),c(-2,0),c(-2,0.9),c(-10,0))
secant_test_results <- sapply(secant_test,secant,f=log_deriv)
data.frame(secant_test_results)

l_prime_theta <- data.frame(theta, likelihood=unlist(lapply(theta, log_deriv)))
ggplot(l_prime_theta)+
  geom_point(mapping=aes(x=theta,y=likelihood)) +
  geom_abline(intercept = 6, slope = -.1,color='red') +
  geom_hline(yintercept=0)+
  labs(title='1st Derivative of Log-likelihood Cauchy(theta,1)',y='log-likelihood')
```
The secant method is particularly relevant when the derivative of the objective
function is exceedingly difficult or impossible to calculate analytically.\
The results again indicate that a good initial guess is crucial. I tested the
function with initial theta1 = 40 and that gave an extremely large number. 
(-3,3) appears to be a worse initial guess than (-2,1) because its resultant
functional value is further from 0. (-10,0) was the worst initial guess given 
500 iterations of the algorithm. This algorithm, namely initial guess (-3,3), 
revealed yet another local maxima at 2.8, further demonstrating the complexity 
of the objective function. \
The secant method can cause some problems because we could have a denominator 
very near or equal to zero, the latter leading to an undefined result. Further,
by looking at the plot we can see that there are several places on the objective
function where the the functional value is the same. If the secant method were
to at any point have an interval with *both corresponding theta values*, the
update method would be 0, and the resultant secant line would not have a root,
causing the algorithm to diverge. Even if the values are not exactly equal but
are close, i.e we take theta_1 = -5 and theta_2 = -2, the resultant secant line
has a root but is very far away from the objective, indicating divergence. The
red line (ideally) demonstrates this point, as its root is at x=60, already
diverging significantly from the main area of interest. This would cause the 
interval to get wider rather than shorter, which is the opposite of our goal. \
The "mountainous" shape of the derivative is inherently tricky for the secant 
method to navigate, as it can very easily fall into one of these sorts of 
"traps" and diverge from the solution.

## e.
```{r}
num_experiments <- 100
newt_times <- rep(0,num_experiments)
bis_times <- rep(0,num_experiments)
fixedp_times <- rep(0,num_experiments)
sec_times <- rep(0,num_experiments)
for (i in 1:num_experiments){
  newt_start <- Sys.time()
  nr_call <- Newt_Raph_MLE(0,vectorized_loglike)
  newt_times[i] <- Sys.time() - newt_start

  bis_start <- Sys.time()
  bisection_call <- bisect(c(-1,1),log_deriv)
  bis_times[i] <- Sys.time() - bis_start

  fixedp_start <- Sys.time()
  fixedp_call <- fixed_point(-1,vectorized_loglike,alpha=64)
  fixedp_times[i] <- Sys.time() - fixedp_start

  sec_start <- Sys.time()
  secant_call <- secant(c(-2,0),log_deriv)
  sec_times[i] <- Sys.time() - sec_start
}
newt_mean <- mean(newt_times)
bis_mean <- mean(bis_times)
fixedp_mean <- mean(fixedp_times)
sec_mean <- mean(sec_times)
cauchy_time <- data.frame(NR_Avg=newt_mean,Bisection_Avg=bis_mean,
                          Fixed_Point_Avg=fixedp_mean,Secant_Avg=sec_mean)
```
Repeat experiment with $N(\theta,1)$\
```{r}
# I re-wrote this function because the original was pretty hard-coded
Newt_Raph_Norm_MLE <- function(ini_guess, x, max_iter=25, tol=1e-05){
  # ini_guess = initial guess 
  # x = vector of observed values
  old_guess <- ini_guess
  new_guess <- old_guess + sum(x-old_guess)/length(x)
  iter <- 1
  while (iter<max_iter & sqrt((old_guess-new_guess)^2)>tol){
    old_guess <- new_guess
    new_guess <- old_guess + sum(x-old_guess)/length(x)
    iter <- iter + 1
  }
  list(Initial=ini_guess,Steps=iter,Estimate=new_guess)
}
norm_values <- rnorm(20,-.1922,1)
normal_first_deriv <- function(theta) sum(norm_values-theta)
normtimes1 <- rep(0,length(num_experiments))
normtimes2 <- rep(0,length(num_experiments))
normtimes3 <- rep(0,length(num_experiments))
normtimes4 <- rep(0,length(num_experiments))
for (i in 1:num_experiments){
  start1 <- Sys.time()
  Newt_Raph_Norm_MLE(4,norm_values)
  normtimes1[i] <- Sys.time() - start1
  
  start2 <- Sys.time()
  bisect(c(0,6),normal_first_deriv)
  normtimes2[i] <- Sys.time() - start2
  
  start3 <- Sys.time()
  fixed_point(5,normal_first_deriv)
  normtimes3[i] <- Sys.time() - start3
  
  start4 <- Sys.time()
  secant(c(-1,1),normal_first_deriv)
  normtimes4[i] <- Sys.time() - start1
}
norm_newt_mean <- mean(normtimes1)
norm_bis_mean <- mean(normtimes2)
norm_fixed_mean <- mean(normtimes3)
norm_sec_mean <- mean(normtimes4)
norm_time <- data.frame(NR_Avg=norm_newt_mean,Bisection_Avg=norm_bis_mean,
                          Fixed_Point_Avg=norm_fixed_mean,Secant_Avg=norm_sec_mean)
rbind(Cauchy=cauchy_time,Normal=norm_time)
```
Newton-Raphson and fixed point both run significantly faster for the Normal 
density than for Cauchy, but bisection runs at half the speed for Normal, which
also sees the secant method performing much worse comparitively. Each method 
seems to have run fairly quickly, but the speed scales with problem complexity, 
and the run-time will vary depending upon initial value specification. Good 
initial value specification for the secant method in Cauchy is likely what led
to its much faster performance.  \

In theory, Newton-Raphson is the fastest provided it is given a solid initial 
guess (though good initial guesses are necessary for the optimal performance of 
all methods). This result is confirmed by the above output for Normal, but not
for Cauchy, for the latter it is the slowest method. NR is the second fastest 
for Normal, though it is very closely behind fixed point. \

It is notable that Newton-Raphson can easily diverge, where an extreme initial 
guess can lead to increasingly worse estimates each iteration. This can be seen 
since guesses -11 and 38 failed to even run for part a). Thus, its stability
is largely dependent upon the quality of the initial guess but also the quality
of the derivative approximations (where exact forms are not available).\

Bisection may be the most stable, as it requires very poor initial guesses to 
not converge, whereas fixed point and secant can diverge easily. That said, our
specific problem is tricky because the derivative changes signs numerous times 
(as seen in part d)). This decreases the stability of both bisection and secant.
Secant operates best when the objective function is "fairly concave" (i.e. 
several local maxima are okay, but the closer they are together the more 
probable the secant method is to diverge). Our function has several local maxima
that are relatively close together, so a combination of theta estimates leading
to similar functional values causes divergence (again illustrated by d)). This
happens because the resultant secant line has an x-intercept very far away from
the region of interest. Despite this, secant performed well for Cauchy, even 
though the derivative has the aforementioned tricky form. This is only because
the initial guesses were solid.\

For fixed point, a poor initial estimate can quickly lead to the algorithm 
spiraling out of control, and it is dependent upon the Lipschitz condition 
holding. In other words, performance depends on a good formulation of G. \

Note: The obvious estimate for the normal example would be the sample mean (the 
UMVUE for $\theta$ in this case), I used the most common optimal value obtained 
by the methods for the Cauchy distribution, as it seemed that is what the 
problem was suggesting.\

# Problem 2.2\
Density: $f(x)=\frac{1-cos(x-\theta)}{2\pi}$\
Likelihood: $L(\theta)=\prod_{i=1}^{n}\frac{1-cos(x_i-\theta)}{2\pi}$\
Log-Likelihood: $l(\theta)=-nlog(2\pi)+\sum_{i=1}^{n}log(1-cos(x_i-\theta))$\
Log-Likelihood with our data (for visualization purposes): \
$l(\theta)=-nlog(2\pi)+log(1-cos(3.91-\theta))+ ... +log(1-cos(2.5-\theta))$\
Deriving wrt theta, checking assuming one x value\
$\frac{1}{1-cos(x-\theta)}*sin(x-\theta)*-1=\frac{-sin(x-\theta)}{1-cos(x-\theta)}$\
First Derivative log-like:$l'(\theta)=\frac{-\sum_{i=1}^{n}sin(x_i-\theta)}{\sum_{i=1}^{n}(1-cos(x_i-\theta))}$\

## a
```{r,warning=FALSE}
x <- c(3.91,4.85,2.28,4.06,3.7,4.04,5.46,3.53,2.28,1.96,
          2.53,3.88,2.22,3.47,4.82,2.46,2.99,2.54,0.52,2.5)
param_space <- seq(-pi,pi,2*pi/199)
log_likelihood2 <- function(theta) {-length(x)*log(2*pi)+sum(log(1-cos(x-theta)))}
# call above function with theta and observed values
plot(param_space, unlist(lapply(param_space, log_likelihood2)),
     xlab='theta',ylab='log-likelihood', main='Log-likelihood for Problem 2.2')
```

## b
Method of moments estimation
\begin{equation}
\begin{split}
E(x) & =\int_{0}^{2\pi}\frac{x(1-cox(x-\theta))}{2\pi}dx\\
&=\frac{1}{2\pi}\int_{0}^{2\pi}(x-xcox(x-\theta))dx\\
&=\frac{1}{2\pi}[\int_{0}^{2\pi}xdx-\int_{0}^{2\pi}xcox(x-\theta))dx]\\
&=\frac{1}{2\pi}[2\pi^2-\int_{0}^{2\pi}xcox(x-\theta))dx]\\
&=\frac{1}{2\pi}[2\pi^2-[xsin(x-\theta)+cos(x-\theta)]_0^{2\pi}]\\
&=\frac{1}{2\pi}[2\pi^2-[2\pi sin(2\pi-\theta)+cos(2\pi-\theta)+cos(0-\theta)]]\\
&=\frac{1}{2\pi}(2\pi^2-2\pi sin(\theta))\\
&=\pi+sin\theta\
\end{split} \end{equation} \
By WLLN, $\bar{x}\rightarrow E(x)$, substitute and solving for $\theta$, we 
have estimator $\hat{\theta}=arcsin(\bar{x}-\pi)$

## c
```{r}
# include method of moments estimator
mom_estimator <- asin(mean(x)-pi)
test_values2 <- c(-2.7,2.7,0,mom_estimator)
results_df2 <- data.frame()
# using same method as previously for viewing results
for (i in 1:length(test_values2)){
  newt_raph_results2 <- Newt_Raph_MLE(test_values2[i],log_likelihood2)
  results_df2[i,1] <- newt_raph_results2[1]
  results_df2[i,2] <- newt_raph_results2[2]
  results_df2[i,3] <- newt_raph_results2[3]
}
results_df2
```

## d
```{r}
results_df3 <- data.frame()
results_df3 <- t(data.frame(sapply(param_space,Newt_Raph_MLE,f=log_likelihood2)))
# all results
head(results_df3,20)
# simplify estimates by rounding
raw_estimates <- unlist(results_df3[,3])
prob2_estimates <- round(raw_estimates,8)
# extract unique estimates for theta
unique_estimates <- unique(prob2_estimates)
# determine where cutoffs for intervals are
indices<- match(unique_estimates,prob2_estimates)
# subset initial guesses according to indices
groups_least_values <- results_df3$Initial[indices]

intervals <- list()
for (i in 1:length(unique_estimates)){
  intervals[[i]] <- unlist(subset(results_df3,round(raw_estimates,8)==unique_estimates[i])[,2])
}
# 19 groups, first three displayed 
head(intervals,3)
```
When rounding to 8 decimal places, we have 19 unique estimates across the entire
parameter space. Vector "groups_least_values" contains the "first" value in the 
initial set to give a different estimate. That is, the entry just before -2.7942
in the param_space vector would give a different final result. Thus, loose
intervals could be constructed from this vector, where any given interval is
defined by the current entry in groups_least_values, up to but not including the
next entry in groups_least_values. The spacing is determined by $2\pi/199$, 
which was used to create the parameter space and also the plot in part a). \
For instance, the first set would be 
$(-\pi,-3.110019,-3.078445,...,-2.825855)$, which all give an estimate of -3.093
The next value in the param_space vector gives an estimate of -2.786167, and the
pattern repeats. Notably, not every set is of the same size.\
The "intervals" object is a list that contains sets for the initial values that
lead to unique final estimates. Any initial value in any particular set would
result in the so-called local modes, which are in turn given by the vector 
"unique_estimates".

## e
My thinking for this problem was to use the end values of the intervals found in
the previous problem. I searched the parameter space near these values by 
gradually incrementing them by some "eps" > 0, looping until the estimate 
obtained by the Newton-Raphson method was different. "Different" constitutes the 
estimates being different when rounded to the 6th decimal place. Further 
precision can be obtained by decreasing the value of eps or rounding to a later
decimal place. It is up to the user to decide what defines the estimates being
"close", hence more insight could be gleaned by less rounding.
```{r}
distinct_results <- function(initial_guess,eps=1e-03){
  # parameters: initial guess, vector of values
  # eps determines which 10s place to nudge the initial guess by
  results1 <- Newt_Raph_MLE(initial_guess,log_likelihood2)
  estimate1 <- round(as.numeric(results1[3]),6)
  # nudge in positive direction
  i <- 1
  nudge <- i*eps
  guess_nudge <- initial_guess + nudge
  results2 <- Newt_Raph_MLE(guess_nudge,log_likelihood2)
  estimate2 <- round(as.numeric(results2[3]),6)
  # loop until the estimates are different
  while(estimate1 == estimate2){
    results1 <- Newt_Raph_MLE(initial_guess,log_likelihood2)
    estimate1 <- round(as.numeric(results1[3]),6)
    # increase size of nudge
    i <- i+1 
    nudge <- i*eps
    guess_nudge <- initial_guess + nudge
    results2 <- Newt_Raph_MLE(guess_nudge,log_likelihood2)
    estimate2 <- round(as.numeric(results2[3]),6)
  }
  initial_val1 <- as.numeric(results1[2])
  initial_val2 <- as.numeric(results2[2])
  list(Diff_Between_Initial=abs(initial_val1-initial_val2),
       Ini_1=initial_val1,Final_1=estimate1,
       Ini_2 = initial_val2,Final_2=estimate2)
}
# Vector of values near threshold, i.e. start with one value, add incrementally
# until a different final result is obtained
# If we test every "close" value, we effectively search the entire space for
# the two values that are closest together but lead to different final estimates
close_values <- unlist(results_df3[,2][indices-1])
differences <- t(sapply(close_values,distinct_results))
differences_mat <- data.frame(differences)
differences_mat
```
Starting values of -2.22593 and -2.222953 have a difference of 0.003 and lead to 
final guesses of -2.232167 and -1.658283, respectively. These initial starting 
values should be the two closest that lead to different estimates *without*
*testing for smaller epsilon.* If we choose a smaller increment, we may discover
even closer values that go to different end results. However, the computational
complexity increases exponentially with smaller epsilon. In the interest of time
I chose a relatively small increment size. 

```{r,include=F}
# I've included this as an appendix, it was my initial method for solving #1
# Following was my writing out of the first derivative of the log-likelihood
Newt_Raph_MLE_orig <- function(ini_guess, max_iter=25, tol=1e-05){
  # ini_guess = initial guess 
  # x = vector of observed values
  old_guess <- as.numeric(ini_guess)
  new_guess <- old_guess - log_deriv(old_guess)/log_deriv_sec(old_guess)
  iter <- 1
  while (iter<max_iter & sqrt((old_guess-new_guess)^2)>tol){
    old_guess <- as.numeric(new_guess)
    new_guess <- old_guess - log_deriv(old_guess)/log_deriv_sec(old_guess)
    iter <- iter + 1
  }
  list(Initial=ini_guess,Steps=iter,Estimate=new_guess)
}

# old functions: log-likelihood written in analytical form, its second derivative
ll1 <- function(x,theta) -length(x)*log(pi)-sum(log(1+(x-theta)^2)) 
log_deriv_sec <- function(theta) 2*sum(((values-theta)^2-1)/(1+(values-theta)^2)^2)

# First derivative log-like: $\sum_{i=1}^{n}\frac{-1}{1+(x-\theta)^2}*2(x-\theta)*-1)$
# $=2\sum_{i=1}^n\frac{x-\theta}{1+(x-\theta)^2}$
# Second derivative log-like: did not obtain "confidently" 

# 2d old commentary
# Rounding to 8 decimal places, my code determined there are 43 unique estimates,
# implying there are 43 groups along the sub-interval. Each entry in the final 
# vector gives the lowest possible value of each respective interval, their 
# subsequent values representing the non-included largest possible value for the 
# interval. Any initial estimates within this interval are expected to give the 
# same final estimate for $\theta$. For example, any initial estimate between
# -2.8890023 and -2.6679857 is expected to give a final result of -2.752389.
# In general, this demonstrates not only the complexity of the objective function,
# but also that there are *many* local maxima. This is to be expected though, as
# the density is periodic. It is probably worth mentioning that each initial 
# estimate leads to a final guess that is very close to it, meaning the algorithm
# does not move far away from the initial guess at any point.
```