---
title: "Stat 705 HW6"
author: "Alex Caramanico"
date: "`r Sys.Date()`"
output: pdf_document
---
I have tried my absolute best with this assignment, I simply ran out of time.
I believe 3.2 is working correctly, but there are certainly ways of 
incorporating the aspiration and diversification criterion more completely. 
For 3.8 Simulated Annealing, the results seem to be working correctly despite
a strange shape in the plot. This is likely because the initial fit is the one
given by the data set, while the consecutive iterations with the same WSS is due
to the same model fit occurring repeatedly. 
I have no idea how the genetic algorithm is supposed to work with this problem,
and for the third problem I have spent hours confused by the proper way to 
update the parameters. The code runs, but does not give logical results. I feel
if we were given more time I'd have been able to produce much more desirable 
(and more cleanly coded) results.

# Problem 3.2
Implement a tabu algorithm for minimizing the AIC for the baseball salary 
regression problem. Model your algorithm after Example 3.7.

## a., b., and c.
Compare the effect of using different tabu tenures. \
Monitor changes in AIC from one move to the next. Define a new attribute that
signals when the AIC change exceeds some value. Allow this attribute to be 
included on the tabu list, to promote search diversity. (indicated via comment)\
Implement aspiration by influence, over-riding the tabu of reversing a low-
influence move if a high-influence move is made prior to the reversal. Measure 
influence with changes in $R^2$. (indicated via comment)
```{r,warning=FALSE,message=FALSE}
# load libraries, set seed
library(tidyverse)
library(readr)
library(ggplot2)
set.seed(1726)
# import data, dataframe set-up
baseball_dat <- read_table('baseball.dat')
baseball_dat$freeagent <- as.factor(baseball_dat$freeagent)
baseball_dat$arbitration <- as.factor(baseball_dat$arbitration)

# THE DIVERSE AND ASPIRE PARAMETERS ARE BOOLEANS, AND CAN BE TURNED ON/OFF TO
# ALTER WHICH COMPONENTS OF THE ALGORITHM ARE RUN
tabu_func <- function(tabu_term,data,iter=75,improvement=0.4,diverse=T,aspire=T){
  predictors <- data[,-1] # REMOVE TARGET VARIABLE
  log_sal <- log(data$salary)

  n_obs <- length(log_sal) # n 
  n_predictors <- length(predictors[1,]) # m

  tabu <- rep(0,n_predictors)
  aic_scores <- rep(0, iter+1) # track AIC values
  aic_diff <- rep(0,iter+1)
  
  better <- NULL
  r2_values <- NULL
  # for b.
  if (diverse){better <- rep(0,iter+1)} 
  # for c.
  if (aspire){r2_values <- rep(0,iter+1)}

  run <- rbinom(n_predictors,1,0.5) # 0 or 1, exclude or include predictor in reg
  run_vars <- predictors[,run==1] # select only variables randomly chosen by binom
  model <- lm(log_sal~.,run_vars) # fit model with the variables
  run_aic <- extractAIC(model)[2]
  best_aic <- run_aic # initialize: first and only run would have the best AIC 
  aic_scores[1] <- run_aic # update vector of AIC scores
  if (aspire) {r2_values[1] <- summary(model)$r.squared} 

  for (i in 1:iter){
    run_aic <- 0
    # TESTS ALL MODELS IN THE 1-NEIGHBORHOOD, CHOOSES THE BEST MODEL IF IT IS 
    # NOT TABU, OTHERWISE SELECT LEAST UNFAVORABLE UNLESS THE MODEL IS THE BEST
    for (j in 1:n_predictors){
      run_step <- run
		  run_step[j] <- !run[j]
		  run_vars <- predictors[,run_step==1]
		  model <- lm(log_sal~.,run_vars)
		  run_step_aic <- extractAIC(model)[2]
		  if(run_step_aic < run_aic && tabu[j]==0){
		    run_next <- run_step
		    run_aic <- run_step_aic
		    pos <- j
		  }
	    if(run_step_aic < run_aic && tabu[j]!=0 && run_step_aic < best_aic){
	      run_next <- run_step
	      run_aic <- run_step_aic
	      pos <- j
	    }
	    # DECREMENT TABU TERMS
	    if(tabu[j]!=0){tabu[j]<-tabu[j]-1}
		    }
	  tabu[pos] <- tabu_term
	  run_current <- run_next
	  if(run_aic < best_aic){
		  best_aic <- run_aic # save best AIC
		  run <- run_current
	  }
	  aic_scores[i+1] <- run_aic
	  
	  # for b: negative -> worse, positive -> better
	  if (diverse) {aic_diff[i] <- aic_scores[i]-run_aic
	    # if "significant" improvement, indicate "Large"
	    if (aic_diff[i] > improvement) {better[i]<-'Large'} 
	    else {better[i]<-'Small'}
	  }
	  # for c:
	  if (aspire){
  	  r2_values[i+1] <- summary(model)$r.squared
	  }
	} 
  list(aic_scores,best_aic,tabu_term,aic_diff,better,r2_values)
}
test_tenures <- c(1,5,10,12,15,20) # weird with 50, 100, aics not changed from 0
results <- sapply(test_tenures, tabu_func,data=baseball_dat)

# setting up dataframe for ggplot
tenure_length <- matrix(sapply(test_tenures,rep,times=76),ncol=1) 
iter_number <- rep(1:76,length(test_tenures))
aic_value <- c(unlist(results[1,1]),unlist(results[1,2]),unlist(results[1,3]),
               unlist(results[1,4]),unlist(results[1,5]),unlist(results[1,6]))
improvements <- c(unlist(results[5,1]),unlist(results[5,2]),unlist(results[5,3]),
               unlist(results[5,4]),unlist(results[5,5]),unlist(results[5,6]))
all_aics <- data.frame(iter_number,tenure_length,aic_value,improvements)

ggplot() + 
  geom_line(data=all_aics, 
            mapping=aes(x=iter_number, y=aic_value,color=as.factor(tenure_length)))+
  geom_point(data=all_aics[all_aics$improvements=='Large',],
             mapping=aes(x=iter_number,y=aic_value,color=as.factor(improvements),shape='.'))+
  ylim(c(-400,-425)) +
  ggtitle('AIC plot per iteration, grouped by tenure length')

# plot best AICs against tenure length
plot(test_tenures,unlist(results[2,]),xlab='Tenure Length',ylab='AIC',
     main='Best AIC versus Tabu Tenure Length')
results[6,] # R^2 values
```
There seems to be some sort of balancing act necessary for the tenure length. If
the length is too long, the algorithm "breaks" and "explores too much" (the code
had several AICs of 0 with tenure lengths of 50 and 100, which were excluded), 
while if the length if too short, it cycles within a single neighborhood of the 
function.The differences between best AICs do not seem very significant, though.
We see in the first plot that tenure lengths 1 and 10 are the best, but 10 is 
able to jump 1. This is likely due to 1 being stuck in a local neighborhood and
not finding an area of the optimization function which is slightly higher.\

Experimentation revealed that the difference difference between the "best best" 
and "worst best" AIC scores is no greater than 10 in absolute value, and since 
the AICs are all in in range of 100s, such a difference is not intuitively 
large. The difference between AICs of say -418.94 and -418.24 may be noteworthy 
depending on the specific problem. The level of precision desired is of course a 
huge factor in this. The second plot shows again that lengths of 1 and 10 are 
the best, while 20 is clearly too long, and 5, 12, 15 are sub-optimal choices. 

# Problem 3.8
Thirteen chemical measurements were carried out on each of 178 wines from three
regions of Italy [53]. These data are available from the website for this book. 
Implement simulated annealing and the genetic algorithm, partition the wines 
into three groups for which the total of the within-group sum of squares is 
minimal. Comment on your work and the results. This is a search problem of size 
$3^p$ where $p = 178$. If you have access to standard cluster analysis routines, 
check your results using a standard method like that of Hartigan and Wong [317].

## Simulated Annealing
```{r,message=FALSE}
wine_data <- read_table('wine.dat')


# computes total within sum of square errors
compute_wss <- function(data){
  region1 <- subset(data,data[,1]==1)
  errs1 <- region1 - colMeans(region1)[col(region1)]
  wss1 <- sum(errs1^2)
  region2 <- subset(data,data[,1]==2)
  errs2 <- region2 - colMeans(region2)[col(region2)]
  wss2 <- sum(errs2^2)
  region3 <- subset(data,data[,1]==3)
  errs3 <- region3 - colMeans(region3)[col(region3)]
  wss3 <- sum(errs3^2)
  return(wss1+wss2+wss3)
}

sim_annealing_cluster <- function(data,temps,tau_start=5,restarts=15,groups=3,alpha=0.9){
  ## INITIAL VALUES
  set.seed(1999) # fix results
  num_reps <- restarts / length(temps) 
  cooling <- c(rep(temps[1],num_reps),rep(temps[2],num_reps),rep(temps[3],num_reps))
  tau <- rep(tau_start,restarts) # 15 random restarts 
  # withinss based on groups given by data
  run_wss <- compute_wss(data) 
  best_wss <- run_wss
  wss <- run_wss
  
  for(j in 2:restarts){tau[j] <- alpha*tau[j-1]} # 10, 9, 8.1, etc
  ## MAIN
  for(j in 1:restarts){
	  for(i in 1:cooling[j]){
	    # randomly select data point, reassign its cluster
	    new_data <- data
	    row <- sample(1:nrow(new_data),1)
	    random_obs <- new_data[row,]
	    random_obs[,1] <- c(new_data[,1]$region) %>%
	                        setdiff(sample(groups,1)) %>% 
	                        sample(1)
	    new_data[row,] <- random_obs
	    run_step_wss <- compute_wss(new_data)
		  
	    p <- min(1,exp((run_wss-run_step_wss)/tau[j])) 
 		  # accept new data (with reassigned row) with probability p
		  if(rbinom(1,1,p)){  
			  data <- new_data
		    run_wss <- run_step_wss}
		  if(run_step_wss < best_wss){
			  data <- new_data 
			  best_wss <- run_step_wss}
		  wss <- c(wss,run_wss) # update vector of sum of squares
	  }
  }
  list(WithinSSError=wss,Best_Predictors=data[,1],Best_error=best_wss)
}
temps <- c(60,120,220)
sim_ann_results <- sim_annealing_cluster(wine_data,temps)
plot(1:length(sim_ann_results[[1]]),sim_ann_results[[1]], pch=20,
     xlab='Iteration',ylab='Total WSS',
     main=sim_ann_results$Best_error)
```
Since we're starting with the initial data set, it takes a while for the 
algorithm to find a better (albeit slightly) grouping of the observations. The 
consecutive values in the plot are due to the algorithm fitting a model with the
exact same groupings, leading to the same total within ss errors value. 
Eventually, we see the algorithm reach lower values than the initial fit.

## Genetic Algorithm
```{r}
genetic_algo_cluster <- function(data,iter=100,m.rate=0.1,P=20,ngroups=3){
  n <- length(data)
  m <- length(data[1,])
  r <- matrix(0,P,1)
  phi <- matrix(0,P,1)
  runs <- matrix(0,P,m)
  runs_next <- matrix(0,P,m)
  runs_wss <- matrix(0,P,1)
  all_wss <- matrix(0,P,iter)
  run <- NULL
  best_wss <- 0
  best_wss_gen <- rep(0,iter)

  # INITIALIZES STARTING GENERATION, FITNESS VALUES
  for(i in 1:P){
    new_data <- data
	  row <- sample(1:nrow(new_data),1)
	  random_obs <- new_data[row,]
	  random_obs[,1] <- c(new_data[,1]$region) %>%
	                        setdiff(sample(groups,1)) %>% 
	                        sample(1)
	  new_data[row,] <- random_obs
	  run_step_wss <- compute_wss(new_data)
	  all_wss[i,1] <- runs_wss[i]
	  if(runs_wss[i] < best_wss){
		  run <- runs[i,]
		  best_wss <- runs_wss[i]
	  }
  }
  r <- rank(-runs_wss)
  phi <- 2*r/(P*(P+1))
  best_wss_gen[1] <- best_wss

  ## MAIN
  for(j in 1:iter-1){
	  # BUILDS THE NEW GENERATION, SELECTING FIRST PARENT BASED ON
	  # FITNESS AND THE SECOND PARENT AT RANDOM
	  for(i in 1:10){
		  parent.1 <- runs[sample(1:P,1,prob=phi),]
		  parent.2 <- runs[sample(1:P,1),]
		  pos <- sample(1:(m-1),1)
		  mutate <- rbinom(m,1,m.rate)
		  runs_next[i,] <- c(parent.1[1:pos],parent.2[(pos+1):m])
		  runs_next[i,] <- (runs_next[i,]+mutate)*2
		  mutate <- rbinom(m,1,m.rate)
		  runs_next[P+1-i,] <- c(parent.2[1:pos],parent.1[(pos+1):m])
		  runs_next[P+1-i,] <- (runs_next[P+1-i,]+mutate)*2
	  }
	  runs <- runs_next

  	# UPDATES AIC VALUES, FITNESS VALUES FOR NEW GENERATION
	  for(i in 1:P){
      new_data <- data
	    row <- sample(1:nrow(new_data),1)
	    random_obs <- new_data[row,]
	    random_obs[,1] <- c(new_data[,1]$region) %>%
	                        setdiff(sample(groups,1)) %>% 
	                        sample(1)
	    new_data[row,] <- random_obs
	    run_step_wss <- compute_wss(new_data)
	    all_wss[i,1] <- runs_wss[i]
		  if(runs_wss[i] < best_wss){
			  run <- runs[i,]
			  best_wss <- runs_wss[i]
	  	}
	  }
	  best_wss_gen[j+1]<-best_wss
	  r <- rank(-runs_wss)
	  phi <- 2*r/(P*(P+1))
  }
  list(all_wss)
}
# genetic_algo_cluster(wine_data)
# --> this is just outputting all 0s
```


# Third Problem
Implement the basic stochastic gradient descent algorithm for MLE (StGD); 
here we want to maximize the function \
$\ell_i(\theta)=\sum_{i=1}^n\ell(\theta;X_i)$ \
($\ell(\theta;X_i)$ is the standard log-likelihood with a single data point Xi) \
Need to compute gradients of each $\ell_i(\theta)$ to compute the updates in SGD
(rather than using all the X's), so that f and the (optional depending on how 
you input f) data inputs should be written accordingly. \
Test code on the Probit regression example from class notes. Step-size 
constraints: Decreasing gains with $\sum_t\gamma_t^2<\infty, \sum_t\gamma_t=\infty$ \
Best convergence speed under sufficient conditions: $\gamma_t\sim t^{-1}$

\
Probit regression: $P(Y=1|X)=\Phi(X^T\beta)$ \
Assume true slopes and intercept are $\beta=(0.5,0.4,-0.3,0.7)$, and $-2$.\
Stochastic gradient descent algorithm: \
$w_{t+1} = w_t - \gamma_t\nabla_wQ(z_t, w_t)$
```{r}
set.seed(1984)
# specify true values
b <- c(0.5,0.4,-0.3,0.7) 
intercept <- -2

# randomly generate data
x <- matrix(rbinom(1600,1,0.5),ncol=4) 

# randomly generate response variable, with probabilities according to probit
response_Prbt <- rbinom(400,1,pnorm(intercept + c(x %*% b))) 

binregLik<-function(betas, int, x, response, dist_func = pnorm){
  prob <- dist_func(c(x %*% betas) + int)
  log(ifelse(response == 1, prob, 1 - prob))
}

compute_gradient <- function(b,int,x,y){
  gradient <- matrix(c(y*x[1]/(x%*%b+int) + x[1]*(1-y)/(1-(x%*%b+int)),
                       y*x[2]/(x%*%b+int) + x[2]*(1-y)/(1-(x%*%b+int)),
                       y*x[3]/(x%*%b+int) + x[3]*(1-y)/(1-(x%*%b+int)),
                       y*x[4]/(x%*%b+int) + x[4]*(1-y)/(1-(x%*%b+int)),
                       y/(x%*%b+int) + (1-y)/(1-(x%*%b+int))),ncol=1)
}

Gradmat<-function(x, f, eps = 1e-08){
  # Calculate the difference-quotient approx gradient (matrix) of an arbitrary 
  # input (vector) function "f" at "x"
  dd = length(x)
  aa = length(f(x))
  epsmat = (diag(dd) * eps)/2
  gmat = array(0, dim = c(aa, dd))
  for(i in 1:dd){
    gmat[, i] <- (f(x + epsmat[, i]) -
    f(x - epsmat[, i]))/eps}
  if(aa > 1) gmat else c(gmat)
}

StGD_probit <- function(guess, f, data, step=0.1, max.iter=100){
  # guess: vector of betas and intercept
  # f is the likelihood for each data point -> binregLik
  # data is the randomly-generated matrix, with rows to be selected at random
  num_vars <- length(data[1,])
  num_obs <- nrow(data)
  
  # initial intercept (bias) and betas (weights)
  intercept <- guess[length(guess)]
  betas <- guess[1:length(guess)-1] 
  for (k in 1:max.iter) { # ***CHANGE TO max.iter WHEN WORKING ***
    for (i in 1:num_obs) { # *** CHANGE TO num_obs WHEN WORKING ***
      # randomly select data point and corresponding response
      index <- sample(1:num_obs, 1)
      row <- x[index,]
      # print(row)
      # y_true <- response[index]
      
      # compute loss (negative log-likelihood) and gradient
      # z = X^T B + intercept
      z <- sum(betas*row) + intercept
      y_pred <- rbinom(1,1,pnorm(z))
      # print(y_pred)
      gradient <- compute_gradient(betas,intercept,row,y_pred)
      # gradient <- Gradmat(row,f) # **couldn't get this to work
      # print(gradient)
      
      # Update weights and bias using gradients and step size
      # --> PLUS because the algorithm minimizes by default, to maximize, negate 
      # the gradient, which changes the operator
      # print(betas)
      betas <- betas + step*gradient[1:length(gradient)-1]
      # print(betas)
      intercept <- intercept + step*gradient[length(gradient)]
      
      step <- 1/i 
    }
  }
  list(Betas = betas, Intercept = intercept)
}
# try initial guess B = (0.3,0.2,-0.1,0.8), with intercept -1
StGD_probit(c(0.3,0.2,-0.1,0.8,-1),binregLik,x)
```