---
title: "Stat 705 HW9"
author: "Alex Caramanico"
date: "`r Sys.Date()`"
output: pdf_document
---
# Load Libraries
```{r,warning=FALSE,message=FALSE}
library(tidyverse) # for pipe operator, data frame manipulation
library(dplyr)
library(dbplyr)
library(ggplot2)
library(bayestestR) # for credible interval
```

# Problem 6.3
Consider finding $\sigma^2=E(X^2)$ when $X$ has the density that is proportional
to $q(x)=e^{-|x|^3/3}$ *All results tabulated at the end of this section.*\

## a.
Estimate $\sigma^2$ using importance sampling with standardized weights. \
```{r,message=FALSE,warning=FALSE}
###  Target Function
q_prop <- function(x) exp(-abs(x)^3/3)
# Calculate normalizing constant
norm_const <- integrate(q_prop,lower=-Inf,upper=Inf)$value
q <- function(x) exp(-abs(x)^3/3) / norm_const
xseq <- seq(-2.5,2.5,0.1)
plot(xseq,q(xseq)) # how to bound this?
 
g <- function(x) dnorm(x)
lines(xseq,g(xseq),col="red") 

n <- 10000
iter <- 1000
variance_estimates1 <- rep(0,iter)
for (i in 1:iter){
  samples <- rnorm(n, mean = 0, sd = 1) # proposal: N(0,1)
  weights <- sapply(samples, function(theta) q(theta) / g(theta)) # calculate weights

  weights_normalized <- weights / sum(weights)
  variance_IS <- sum(samples^2 * weights_normalized) # we know true mean is 0
  variance_estimates1[i] <- variance_IS
}
```

## b.
Repeat the estimation using rejection sampling. \
```{r}
M <- (1/norm_const) / (dnorm(0)) # sup (q(x)/g(x))
variance_estimates2 <- rep(0,iter)
for (i in 1:iter){
  Y <- rnorm(n) # sample of size n from envelope g
  gvals <- g(Y) # evaluate envelope at Y
  fvals <- q(Y) # evaluate target at Y
  U <- runif(n)
  keep_inds <- U < fvals/(gvals * M) # rejection sampling algorithm
  x_kept <- Y[keep_inds]
  variance_estimates2[i] <- var(x_kept)
}
rm(Y,gvals,fvals,U)
hist(x_kept,breaks=50,prob=T,main="Kept Samples from Last Run")
lines(xseq,q(xseq),col="red")
```


```{r}
all_results <- cbind(variance_estimates1,variance_estimates2)
data.frame(Method=c("Rej","Imp"),Mean = colMeans(all_results),Repetitions=c(iter,iter))

all_results <- all_results %>% data.frame() %>% pivot_longer(cols=c(variance_estimates1,variance_estimates2))
ggplot(data = all_results) +
  geom_freqpoly(mapping=aes(x=value,color=name),bins=50) +
  ggtitle("Frequency Polygons for Variance Estimates (LAST ITERATES)")
```

Either of these methods probably could have been done without making q(x) a
proper density, but the trade-off would just be scaling the candidate (normal)
density by some constant C. e.g. if we did not normalize q(x), the candidate 
g(x) would be something like 3*dnorm(x), which is not a valid density but 
operates the same way for the  algorithm, since the algorithm is not dependent 
upon having valid densities. The approach I used may not be as tractable for 
target functions (densities) where the normalizing constant is extremely 
difficult to calculate.

# Problem 6.4
Change-point model for coal-mining disaster data. $j=1,...,112$ for 1851 to 1962.
$X_j$: number of accidents in year $j$ \
$X_1,...,X_\theta \sim iid Poi(\lambda_1)$, $X_{\theta+1},...,X_{112} \sim iid Poi(\lambda_2)$,
$\theta \in (1,...,111)$. Change-point occurs after $\theta$th year. \
Three sets of priors for a Bayesian analysis of this model. In each case, 
consider sampling from the priors as the first step of applying the SIR 
algorithm for simulating from the posterior for the model parameters. Of primary 
interest is inference about $\theta$. Use the SIR approach for the following.\

## a. 
Assume a discrete uniform prior on $\theta$, and priors \
$\lambda_i|a_i \sim Gamma(3,a_i)$ \
$a_i \sim Gamma(10,10)$, independently for $i=1,2$. \
Denote the Gamma density function with $\Gamma(x;shape=\alpha,scale=\beta)$ \
Posterior $\propto$ uniform prior * joint prior * likelihood \
$Posterior \propto P(\theta)*P(\lambda_i,a_i)*f(x_1,...x_n|\lambda_i,a_i)$ \
$\propto P(\theta)*P(\lambda_i|a_i)P(a_i)*f(x_1,...x_n|\lambda_i,a_i)$ \
$\propto \frac{1}{112}*\Gamma(\lambda_1;3,a_1)*\Gamma(a_1;10,10)*\Gamma(\lambda_2;3,a_2)*\Gamma(a_2;10,10)*\prod_{1}^{\theta} Poi(x_i;\lambda_1) * \prod_{\theta+1}^{112} Poi(x_i;\lambda_2)$ \
```{r,message=FALSE}
# set-up, plots
coal <- read_table("C:/Users/jacma/OneDrive/school work/UMD/Stat 705 Computational Stats/Data/datasets/coal.dat")
ggplot(data=coal,mapping=aes(x=year,y=disasters)) +
  geom_line() +
  ggtitle("Coal Disasters per Year")

x <- coal$disasters
# essentially looking for the "elbow" point of the following graph
cumsum(x) %>% plot(main="Cumulative Number of Coal Mining Disasters")
```

### i. 
Estimate the posterior mean for $\theta$, providing a histogram and credible 
interval \
```{r,message=FALSE,warning=FALSE}
set.seed(987)
n <- 10000
n_resample <- 100000

# Sample from prior
theta <- sample(1:(length(x)-1),n,replace=T) # change point can't occur in last year
a1 <- rgamma(n,shape=10,rate=10)
a2 <- rgamma(n,shape=10,rate=10)
lam1 <- rgamma(n,shape=3,rate=a1)
lam2 <- rgamma(n,shape=3,rate=a2)
param_matrix <- cbind(theta,a1,a2,lam1,lam2)
rm(theta,a1,a2,lam1,lam2)

likelihood <- function(index){
  t <- param_matrix[index,1]
  lam1 <- param_matrix[index,4]
  lam2 <- param_matrix[index,5]
  seq1 <- 1:t # 1,2,...,theta
  seq2 <- (t+1):length(x) # theta+1, theta+2, ... 112
  return(exp(sum(dpois(x[seq1],lam1),log=T) + sum(dpois(x[seq2],lam2),log=T)))  
}
# weight by likelihood
sir_weights <- sapply(1:n,likelihood)
sir_weights <- sir_weights / sum(sir_weights)

# resample X1,...,Xn from Y1,..,Ym with replacement with probabilities w(Y1),..., w(Ym)
# i.e. resample the ROWS so we can hold onto each parameter value
resampled <- param_matrix[sample(1:n,n_resample,replace=TRUE,prob=sir_weights),] %>% data.frame()

theta_resamples <- data.frame(change=resampled$theta)
post_mean <- mean(theta_resamples$change)
post_var <- var(theta_resamples$change)
ggplot(data=theta_resamples,mapping=aes(x=change)) +
  geom_histogram(bins=50) +
  geom_vline(xintercept = ci(theta_resamples$change,method="HDI")$CI_low, color = "red") + 
  geom_vline(xintercept = ci(theta_resamples$change,method="HDI")$CI_high, color = "red") + 
  ggtitle("Histogram of Posterior Means (95% Credible Interval in Red)",
          toString(c("Mean:",mean(theta_resamples$change))))
```

### ii. 
Provide similar information for estimating $\lambda_1, \lambda_2$ \
```{r,warning=FALSE}
lambda_resamples <- data.frame(lam1=resampled$lam1,lam2=resampled$lam2)
ggplot(data=pivot_longer(lambda_resamples,cols=c(lam1,lam2)))+
  geom_histogram(mapping=aes(x=value,fill=name),bins=100) +
  xlim(c(0,8)) +
  ggtitle("Histogram of Re-Sampled Lambda 1 and Lambda 2",
          toString(c("Means:",colMeans(lambda_resamples) %>% round(2))))
```
The x-axis limits were truncated since there were only a couple of lambdas 
greater than 8 that were re-sampled. They would not have even been visible on
the plot and were thus omitted. The means suggest that after the 39th year, the
rate of coal mining disasters per year changes from 2.65 to 0.41. $\lambda_2$
appears to be pretty heavily concentrated around its mean, whereas $\lambda_1$
almost appears to be multimodal, with peaks at 0.4, 2, 2.8, and 4. 

### iii. 
Make a scatter plot of $\lambda_1$ against $\lambda_2$ for the initial SIR 
sample, highlighting the points re-sampled at second stage of SIR. \
```{r}
# use ifelse statement: check if initial lambda sample was re-sampled 
# (i.e. in the selected rows)
all_lams <- data.frame(lam1 = param_matrix[,4],lam2 = param_matrix[,5],
                       repicked=ifelse(round(param_matrix[,4],3) %in% round(resampled$lam1,3)
                                       & round(param_matrix[,5],3) %in% round(resampled$lam2,3),
                                       "Yes","No"))

ggplot(data=all_lams) +
  geom_point(mapping=aes(x=lam1,y=lam2,color=repicked),size=0.5) +
  ggtitle("Lambda1 vs Lambda 2, Re-Sampled Points Highlighted")
```
A slight pattern certainly emerges here (though much of the plot is somewhat 
hard to see because of the overlap). We see that almost all of the most extreme
lambda values (in either axis) are not re-sampled - the blue points never 
exceeding about 14 for either lambda 1 or lambda 2. This makes sense, because 
the data set does not have mining disasters occurring at such an extreme rate. 
The most we saw in any year was 6, so a rate parameter of much higher than that
makes no sense. While it can be argued that a parameter around 10-14 is also
illogical, these points are in the vast minority of re-sampled values. That is,
almost all of the re-sampled points are within $\lambda_1\in(0,8), \lambda_2 \in (0,2)$.
This assertion is further supported by the preceding histogram.

### iv.
Report the following: \
  a) initial and re-sampling sample sizes \
  b) the number of unique points and highest observed frequency in re-sample \
  c) a measure of the effective sample size for importance sampling in this case \
```{r}
# b)
theta_occurences <- table(resampled$theta) %>% data.frame()
nrow(theta_occurences)
colnames(theta_occurences) <- c("Theta","Freq")
theta_occurences[which(theta_occurences$Freq==max(theta_occurences$Freq)),]

# c) formula from book pg 173
1/sum(sir_weights^2)
```
The initial sample size was 10,000, while there were 100,000 re-sampled values.
Every value of theta was re-sampled at least once, but the most common was 39 
(corresponding to 1890 being the change point) which was re-sampled 11,406 times. 
The estimate for effective sample size is about 33.


```{r,include=F,echo=F}
# old code doesn't work for plotting because i changed parameters of R functions
g <- function(params) {
  t <- params[1]
  a1 <- params[2]
  a2 <- params[3]
  lam1 <- params[4]
  lam2 <- params[5]
  
  seq1 <- 1:t # 1,2,...,theta
  seq2 <- (t+1):length(x) # theta+1, theta+2, ... 112
  mle_lam1 <- mean(x[seq1]) # use MLE for parameter to create envelope g
  mle_lam2 <- mean(x[seq2]) 
  return((1/112) * dgamma(lam1,shape=3,rate=a1) * dgamma(a1,shape=10,rate=10) * 
                   dgamma(lam2,shape=3,rate=a2) * dgamma(a2,shape=10,rate=10)*
           sum(dpois(x[seq1],mle_lam1,log=TRUE)) * sum(dpois(x[seq2],mle_lam2,log=TRUE))) 
}

# format data for plotting
#all_func_vals <- rbind(cbind(rep("f",111),1:111,sapply(1:111,likelihood)),
#                       cbind(rep("g",111),1:111,sapply(1:111,g))) %>% data.frame()
#colnames(all_func_vals) <- c("Function","Theta","Functional Value")
#all_func_vals$Theta <- as.numeric(all_func_vals$Theta)
#all_func_vals$`Functional Value` <- as.numeric(all_func_vals$`Functional Value`)

#ggplot() +
#  geom_point(data=all_func_vals,mapping=aes(x=Theta,y=`Functional Value`,color=Function)) +
#  ggtitle("Posterior Density (f) and Candidate Function (g) Final Iterate")


# --> mcp library makes this really easy (literally does the whole problem), see https://lindeloev.github.io/mcp/articles/poisson.html
```

