---
title: "EPIB 741 HW3"
author: "Alex Caramanico"
date: "`r Sys.Date()`"
output: pdf_document
---
# Data Prep
```{r,echo=F,message=F,warning=F}
library(ggplot2)
library(tidyverse)
library(haven)
library(cowplot)
library(survival)
library(survminer) 
# USING: https://rpkgs.datanovia.com/survminer/#:~:text=survminer:%20Survival%20Analysis%20and%20Visualization,most%20significant%20relation%20with%20survival.
data1 <- read_sas("C:/Users/jacma/OneDrive/school work/UMD/EPIB 741 Association Models/bladder.sas7bdat")

data1 %>% 
  group_by(PATIENT) %>% 
  summarize(meanTIME=mean(TIME),
            meanSTATUS=mean(STATUS),
            meanTREAT=mean(TREAT),
            meanINIT=mean(INIT),
            meanSIZE=mean(SIZE)
            )

# each patient is duplicated three times... why?
# **specifying id=PATIENT in cox models gives same results as just using data2**

data2 <- unique(data1)
data2 %>% subset(data2$STATUS==1) # recurrences
data2 %>% subset(data2$STATUS!=1) # censored
data2$TIME %>% unique() %>% length() # 28 unique times, so 58 are tied
```

# Part A
## Survival Analysis
### 1. and 2. 
Plot the Kaplan-Meier estimator of the survival function (survival curve) for 
patients receiving thiotepa treatment. a) Estimate the probability of recurrence 
free at 12 months, b) Estimate the median recurrence free survival time and its 
95% confidence interval. \

Test the null hypothesis that there is no overall difference between survival 
curves for the two groups (thiotepa and placebo). Based on the test and the two 
survival curves, state your conclusion from the log-rank test
```{r,echo=F,message=F,warning=F}
### 1 and 2
survival.fit <- survfit(Surv(TIME, STATUS) ~ TREAT, id=PATIENT, data=data1)
ggsurvplot(survival.fit, data=data1, 
           risk.table=T, pval=T, conf.int=F, ncensor.plot=T,        
           xlim=c(0,60), xlab="Time in months", break.time.by=6, 
           ggtheme=theme_light(), risk.table.y.text.col=T, risk.table.y.text=T)

### 1a. recurrence free at 12 months
summary(survival.fit)
# P(recurrence free at 12 months for TREATMENT group) = 58.15%
# P(recurrence free at 12 months for CONTROL group) = 22.4%

### 1b. median recurrence free survival time = 19, 95% CI (12, 33)
print(survival.fit)

# survfit(Surv(TIME, STATUS) ~ TREAT, data=data2) %>% print() # identical to above
```

## Cox Proportional Hazards Model
### 3.
Fit a Cox PH model using the treatment as the only independent variable and 
report HR and its 95% confidence interval (CI). 
```{r,echo=F,message=F,warning=F}
# USING: https://www.sthda.com/english/wiki/cox-proportional-hazards-model
ph.fit1 <- coxph(Surv(TIME, STATUS) ~ TREAT, id=PATIENT, data = data1)
summary(ph.fit1)

coxph(Surv(TIME, STATUS) ~ TREAT, data = data2) %>% summary # slightly different...attenuated
```

### 4. Check the PH assumption
```{r,echo=F,message=F,warning=F}
ggsurvplot(survival.fit, fun = "cloglog", data = data1)
# KS test --- *can't seem to find a quick way to do this in R? unless below is the same*
cox.zph(ph.fit1) %>% print()
# ^ this can be applied to the other PH models but he did not ask for it

ggsurvplot(survfit(Surv(TIME, STATUS) ~ TREAT, id=PATIENT, data=data2), 
           fun = "cloglog", data = data2)
# ^ the exact same as the data with duplicated rows
```
survival curves diverge, despite not crossing. log-log plots not parallel. test
of residuals rejects the null hypothesis, we reject the null hypothesis, so we
have increased concerns about the PH assumption NOT being met. 

# Part B - Assume PH Assumption Met
## 5.
Assess the interaction effect of the treatment and the initial number of tumors
```{r,echo=F,message=F,warning=F}
ph.fit2 <- coxph(Surv(TIME, STATUS) ~ TREAT+TREAT*INIT, id=PATIENT, data = data1)
(ph.results2 <- summary(ph.fit2))

# interaction_effect <- function(covariate,model_results){
  # HR = exp(-1.00074+0.04005*INIT)
  # HR <- exp(model_results$coefficients[1,1]+
    #           model_results$coefficients[1,3]*covariate)
  # return(HR)
# }
# sapply(c(1:8),interaction_effect,model_results=ph.results2)

cbind(`# Initial Tumors`=c(1:8),
      HR=exp(ph.results2$coefficients[1,1]+
              ph.results2$coefficients[1,3]*c(1:8)))
```
insignificant modifying effect, treatment effect became stronger (further from 1, 
implying even greater survival for the treatment group). Hazard ratio increases 
monotonically as # of initial tumors increases, matching intuition. More tumors 
increases likelihood of recurrence. at 4 tumors, patients in treatment begin to 
have a greater risk of recurrence than those in control

## 6. 
Suppose the size of the largest initial tumor is a potential confounder, does it 
actually confound the above interaction effect?
```{r,echo=F,message=F,warning=F}
ph.fit3 <- coxph(Surv(TIME, STATUS) ~ TREAT+TREAT*INIT+SIZE, id=PATIENT, data = data1)
ph.results3 <- summary(ph.fit3)
(ph.results3$coefficients[4,2] - ph.results2$coefficients[3,2]) /
  ph.results2$coefficients[3,2] * 100

rm(survival.fit,ph.fit1,ph.fit2,ph.fit3,ph.results2,ph.results3)
```
change in estimate of about 8.75%, so largest initial tumor is probably not a 
confound of the interaction effect, using a 10% threshold to denote large 
changes


# Part C - Assume PH Assumption Not Met
## 7. 
Use the extended Cox model to estimate the effect of the treatment on the 
survival time, controlling for the size of the largest initial tumor as a 
confounder. Do not try to assess the interaction effect.
```{r,echo=F,message=F,warning=F}
# log log plots appear to diverge between groups around 15 months in
```

## 8. 
Assume that the size of the largest initial tumor violates the PH assumption, 
use the stratified Cox model to estimate the effect of the treatment on the 
survival time.
```{r,echo=F,message=F,warning=F}
ggplot(data1, aes(x = SIZE)) +
  geom_histogram(binwidth = 0.5, fill = "lightblue", color = "black") +
  stat_bin(binwidth = 0.5, aes(label = ifelse(after_stat(count) == 0, NA, 
                                              after_stat(count))),
           geom = "text",vjust = -0.5) +
  labs(x="Diameter of largest initial tumor (cm)",y="Number of occurences",
       title="Distribution of Largest Initial Tumor Sizes")

# where is the cutoff 
survival.fit2 <- survfit(Surv(TIME, STATUS) ~ SIZE, id=PATIENT, data=data1)
ggsurvplot(survival.fit2, data=data1, 
           xlim=c(0,60), xlab="Time in months", break.time.by=6, 
           ggtheme=theme_light())

data1$SIZE_STRATA <- ifelse(data1$SIZE>1,1,0) # size 1 versus all else - most balanced
data1$SIZE_STRATA2 <- ifelse(data1$SIZE>3,1,0) # >3 considered large https://bcan.org/facing-bladder-cancer/bladder-cancer-types-stages-grades/?gad_source=1&gad_campaignid=21909057091&gbraid=0AAAAADQO1ftuEC6EjRTGps14HP3eZihdt&gclid=CjwKCAjw-dfOBhAjEiwAq0RwIwWkQl1avGIQrYpztXmA6xMCdPaDeWC3Wu-Ql2ohx2wTaD7Ui0TyCBoCIoQQAvD_BwE

ph.fit5 <- coxph(Surv(TIME,STATUS)~TREAT+strata(SIZE_STRATA), id=PATIENT, data=data1)
summary(ph.fit5)$coefficients

ph.fit6 <- coxph(Surv(TIME,STATUS)~TREAT+strata(SIZE_STRATA2), id=PATIENT, data=data1)
summary(ph.fit6)$coefficients

rm(survival.fit2,ph.fit5,ph.fit6)
```
Attenuated affect of treatment (increased, became closer to 1, but still 
statistically significant) when stratifying by size of largest initial tumor. 
